---
title: "Impact of SES & perinatal factors on CPR registration of Children: A survival analysis"
author: "Dr E. O. Abubakar"
date: "2026-02-03"
output: html_document
editor_options: 
  markdown: 
    wrap: 85
---

Variables already prepared will be analysed in this document. R Project name: `child_nglct_proper_sdc`

# import packages

```{r}

# install.packages("graphPAF")

pacman::p_load(
      # "AF",
      "MASS",
      "mgcv",
      "assertr",
      "beepr",
      "broom",
      "gtsummary",
      "knitr",
      "kableExtra",
      "mice",
      "janitor",
      "patchwork",
      "qs",
      "rlang",
      "rmarkdown",
      "RODBC",
      "sailr",
      "scales",
      "speedglm"
  )

library(tictoc)
library(sjmisc)
library(SmartEDA)
# easystats::easystats_update("core")
library(easystats)
library(survival)
library(survminer)
library(ggsurvfit)
library(fastDummies)
library(tidyverse)
library(graphPAF)

# for automated correlation of all variable types, beyond continous, and their combinations
# install.packages("smartcor", dependencies = T)
library(smartcor)

library(parallel)
options(boot.parallel="snow")
# use a quater of available
options(boot.ncpus=as.integer(parallelly::availableCores()*0.75))

#make outputs directory (if it doesn't already exist)
if(!dir.exists("outputs")){
  dir.create("outputs")
}

gc()
```

# Load, explore & preprocess relevant data

## Census 11 & 21 (Mother & Child) - Cleaned

```{r}

# load cleaned census data
qload("S:/1409 - Tackling inequalities in child health An interventional epidemiology platfor/Ele/child_nglct/data/mother_chld_census.qsm")

## mother-child in census 2011

mother_child_cens11 <-
  mother_child_cens11 %>% 
  mutate(
    SEX.cohrt = case_match(SEX.cohrt,
                           1 ~ "Male",
                           2 ~ "Female",
                           .default = as.character(SEX.cohrt)
                           ) %>% factor,
    GNDR_CD = case_match(GNDR_CD,
                           1 ~ "Male",
                           2 ~ "Female",
                           .default = as.character(GNDR_CD)
                           ) %>% factor,
    ALF_STS_CD = case_match(ALF_STS_CD,
                            4 ~ "04: Direct match on identifiers",
                            39 ~ "39: Probabilistic match at 90%+",
                            35 ~ "35: Probabilistic match at 50%+",
                            99 ~ "99: No match",
                            NA ~ "NA: Missing info",
                            .default = as.character(ALF_STS_CD)
                            ) %>% factor,
  ) %>% 
  droplevels() %>%
  glimpse

mother_child_cens11 %>%
  frq(GNDR_CD, SEX.cohrt, ALF_STS_CD)

mother_child_cens11 %>% glimpse

## mother-child in census 2021
mother_child_cens21 <-
  mother_child_cens21 %>%
  mutate(
    SEX.mc_cohrt = case_match(SEX.mc_cohrt,
                              1 ~ "Male",
                              2 ~ "Female",
                              .default = as.character(SEX.mc_cohrt)
                              ) %>% factor,
    GNDR_CD = GNDR_CD %>% as.character() %>% case_match(.,
                           "1" ~ "Male",
                           "2" ~ "Female",
                           .default = as.character(GNDR_CD)
                           ) %>% factor,
    across(c(where(is.character), -QUALIFICATIONS, -WELSH_LANGUAGE_PROFICIENCY, -ECONOMIC_INACTIVE_STATUS), factor)
  ) %>% 
  droplevels() %>%
  glimpse

mother_child_cens21 %>%
  frq(GNDR_CD, SEX.mc_cohrt)

```

Focus is on table containing data related to reasons for being on a child protection
plan (i.e. being on the child protection register - CPR)

## CINW (2010 - 2016 ended)

```{r}

# load data on CPR (returns cinw_full, crcs_full, lacw_full)
qload("S:\\1409 - Tackling inequalities in child health An interventional epidemiology platfor\\Ele\\child_nglct\\data\\chld_nglct_join.qsm")

# clean cinw (2010 - 2016 ended)
cinw_full_cln <-
  cinw_full %>% 
  select(-c(ALF_MTCH_PCT, EXTRACT_DT)) %>%
  mutate(
    across(c(CHILD_HEALTH_SURVEILLANCE_CHECKS, CHILD_PROTECTION_REGISTER, starts_with(c("DISABILITY_", "HEALTH_", "PARENTING_CAPACITY_", "RECENT_REFERRAL_")), YOUTH_OFFENDING, LOOKED_AFTER_CHILD, UNBORN_CHILD),
           ~case_match(.x,
                      1~ "True",
                      0 ~ "False",
                      80 ~ "Not applicable",
                      88 ~ "Not available",
                      .default = as.character(.x))
           ),
    GNDR_CD = case_match(GNDR_CD,
                         1 ~ "Male",
                         2 ~ "Female",
                         3 ~ " Non-binary",
                         4 ~ "Unknown",
                         .default = as.character(GNDR_CD)
                         ),
    CATEGORY_OF_NEED_CODE = as.character(CATEGORY_OF_NEED_CODE) %>%
      str_to_upper() %>%
      factor %>% 
      fct_recode(.,
                 "Abuse or neglect"                = "N1",
                 "Disability or illness"           = "N2",
                 "Parental disability or illness"  = "N3",
                 "Family in acute stress"          = "N4",
                 "Family dysfunction"              = "N5",
                 "Socially unacceptable behaviour" = "N6",
                 "Undocumented"                    = "N7",
                 "Absent parenting"                = "N8",
                 "Adoption disruption"             = "N9"
                 ),
    ASYLUM_SEEKER_CODE = case_match(ASYLUM_SEEKER_CODE,
                                    0 ~ "Not an Asylum-Seeking Child",
                                    1 ~ "Unaccompanied Asylum-Seeking Child",
                                    2 ~ "Member of an Asylum-Seeking family",
                                    .default = as.character(ASYLUM_SEEKER_CODE)
                                    ),
    ETHNIC_ORIGIN_CODE  = as.character(ETHNIC_ORIGIN_CODE) %>%
      str_squish() %>% str_to_upper() %>%
      factor %>%
      fct_recode(.,
                                    'White british' = 'WBRI',
                                    'White Irish' = 'WIRI',
                                    'Traveller' = 'WIRT',
                                    'Gypsy/Gypsy Roma' = 'WROM',
                                    'Any other white background' = 'WOTH',
                                    'White & Black carribean' = 'MWBC',
                                    'White & Black African' = 'MWBA',
                                    'White & Asian' = 'MWAS',
                                    'Any other mixed background' = 'MOTH',
                                    'Indian' = 'AIND',
                                    'Pakistani' = 'APKN',
                                    'Bangladeshi' = 'ABAN',
                 'Any other Asian Background' = 'AOTH',
                 'Carribean' = 'BCRB',
                 'African' = 'BAFR',
                 'Any other black background' = 'BOTH',
                 'Chinese or Chinese British' = 'CHNE',
                                    'Other ethnic group' = 'OOTH',
                                    'Refused' = 'REFU',
                                    'Not obtained' = 'NOBT',
                                    ),
    SOURCE_MOST_RECENT_REFERRAL_CODE = fct_recode(SOURCE_MOST_RECENT_REFERRAL_CODE,
                                    'Primary/community health' = 'R1',
                                    'Secondary health' = 'R2',
                                    'Self referral' = 'R3',
                                    'Family, friend or neighbour' = 'R4',
                                    'Central govt agency' = 'R5',
                                    "LA social support dept" = 'R6',
                                    'Independent provider agency' = 'R7',
                                    'LA housing dept or housing association' = 'R8',
                                    'Other depts of own or other LA' = 'R9',
                                    'Police' = 'R10',
                                    'Other agency' = 'R11',
                                    'Other individual' = 'R12',
                                    ),
    ALF_STS_CD = case_match(ALF_STS_CD,
                            4 ~ "04: Direct match on identifiers",
                            39 ~ "39: Probabilistic match at 90%+",
                            35 ~ "35: Probabilistic match at 50%+",
                            99 ~ "99: No match",
                            NA ~ "NA: Missing info",
                            .default = as.character(ALF_STS_CD)
                            ),
    across(where(is.character), factor)
  ) %>% 
  droplevels() %>%
  relocate(ends_with("_PE")) %>% 
  glimpse

```

## CRCR (2017 - 2021 continuing)

```{r}
# clean crcs (2017 - 2021 continuing)
crcs_full_cln <- 
crcs_full %>% 
  select(-c(ALF_MTCH_PCT, EXTRACT_DT)) %>% 
  mutate(
    across(c(CHILD_HEALTH_SURVEILLANCE_CHECKS, CHILD_PROTECTION_REGISTER, starts_with(c("CPR_", "DISABILITY_", "HEALTH_", "PARENTING_CAPACITY_")), -CPR_DATE, YOUTH_OFFENDING, LOOKED_AFTER_CHILD),
           ~case_match(.x,
                      1~ "True",
                      0 ~ "False",
                      80 ~ "Not applicable",
                      88 ~ "Not available",
                      .default = as.character(.x))
           ),
#     For CPR fields, NA means not on a CPR
    across(c(starts_with("CPR_"), -CPR_DATE),
           ~fct_na_value_to_level(.x,
                                  level = "Not applicable")
           ),
    GNDR_CD = case_match(GNDR_CD,
                         1 ~ "Male",
                         2 ~ "Female",
                         3 ~ " Non-binary",
                         4 ~ "Unknown",
                         .default = as.character(GNDR_CD)
                         ),
    ETHNIC_ORIGIN_CODE = fct_recode(ETHNIC_ORIGIN_CODE,
                                    'White' = 'WHTE',
                                    'Mixed ethnic groups' = 'MIXD',
                                    'Asian or Asian British' = 'ASAB',
                                    'Black African/British, or Caribbean' = 'BBAC',
                                    'Other ethnic group' = 'OOTH',
                                    'Refused' = 'REFU',
                                    'Not obtained' = 'NOBT',
                                    ),
    CATEGORY_OF_NEED_CODE = fct_recode(CATEGORY_OF_NEED_CODE,
                                       "Abuse or neglect"                = "N1",
                                       "Disability or illness"           = "N2",
                                       "Parental disability or illness"  = "N3",
                                       "Family in acute stress"          = "N4",
                                       "Family dysfunction"              = "N5",
                                       "Socially unacceptable behaviour" = "N6",
                                       "Undocumented"                    = "N7",
                                       "Absent parenting"                = "N8",
                                       "Adoption disruption"             = "N9"
                                       ),
    ASYLUM_SEEKER_CODE = case_match(ASYLUM_SEEKER_CODE,
                                    0 ~ "Not an Asylum-Seeking Child",
                                    1 ~ "Unaccompanied Asylum-Seeking Child",
                                    2 ~ "Member of an Asylum-Seeking family",
                                    .default = as.character(ASYLUM_SEEKER_CODE)
                                    ),
    ALF_STS_CD = case_match(ALF_STS_CD,
                                4 ~ "04: Direct match on identifiers",
                                39 ~ "39: Probabilistic match at 90%+",
                                35 ~ "35: Probabilistic match at 50%+",
                                99 ~ "99: No match",
                                NA ~ "NA: Missing info",
                                .default = as.character(ALF_STS_CD)
                                ),
    across(where(is.character), factor)
  ) %>% 
  droplevels() %>% 
  relocate(ends_with("_PE")) %>% 
  glimpse

```

## Join CSC cohort for analysis

The cohort is individuals (children) whose birth period is btw Jan 2009 (CINW start
data) and Dec 2011 (census date). THis is a 4 year period. For children born b4 2011,
there mother's HH census attributes would be thesame as theirs.

The study end date would be data at which people born in 2011 (census year), would
have clocked 18, at which time they are no longer under children care services, i.e.
2029. Previous study ended at "2021-03-31"

censorship is when an individual became recorded on the CPR On the CINW and CRCS
tables each person appears multiple times, yearly. I am interested in the first date
(year, based on DATE_OF_DATA_COLLECTION) they became a recorded in the CPR

I join CINW to CRCS focusing on variables common to both data sets.

RQ: What is the risk that a child will be recorded in the CPR register prior to
reaching age 18.

The cohort is individuals born btw 2009/01 and 2011/12. They were followed up until
whichever cam first between their 18th bday and the end of the study by 2021/03

```{r}
# Stack CINW to CRCS retaining only columns common to both tables

crcs_full_cln %>%
  glimpse

cinw_crcs_join <-
bind_rows(
  list(crcs=crcs_full_cln %>% 
  select(any_of(colnames(cinw_full_cln)), CPR_DATE),
  cinw=cinw_full_cln %>% 
  select(any_of(colnames(crcs_full_cln)))),
  .id = "dta_orign"
) %>% #glimpse
#  for cinw-sourced records (and other not on CPR), fill CPR_DATE with DATE_OF_DATA_COLLECTION
  mutate(
    CPR_DATE = coalesce(CPR_DATE, DATE_OF_DATA_COLLECTION),
    DATE_OF_DATA_COLLECTION = NULL
  ) %>% 
  arrange(CHILD_CODE_PE, CPR_DATE) %>% 
  glimpse

# CSC data end date
# 2021-05-31

# trace factors per child
# "DISABILITY_", "HEALTH_", "PARENTING_CAPACITY_", "RECENT_REFERRAL_"
rm(cinw_crcs_cln)
cinw_crcs_cpr <-
cinw_crcs_join %>% 
  select(!starts_with(c("CPR_", "DISABILITY_", "HEALTH_", "PARENTING_CAPACITY_" )), -c(CHILD_HEALTH_SURVEILLANCE_CHECKS), CPR_DATE) %>% #glimpse
  arrange(CHILD_CODE_PE, CPR_DATE) %>%
  group_by(CHILD_CODE_PE) %>% 
  mutate(
    ttl_rcrds = n(),
    rcrd_rnk = dense_rank(CPR_DATE)
  ) %>% #glimpse
  #   get children on the CPR
  filter(CHILD_PROTECTION_REGISTER == "True") %>%
#   first date on CPR
  # filter(min(DATE_OF_DATA_COLLECTION)) %>% glimpse
  slice_head(n = 1) %>% #glimpse
  ungroup() %>%
    # compute time to first record on CPR
#   compute CPR record to Study end & CPR record to 18 years
#   take whichever one comes first as the end point
  mutate(
#     time of first cpr to end of CSC services, indeed the oldest person in the cohort (born btw 2009 and 2011) will be 12yrs at the time of study end
    # c_18yrs_dt = WOB + years(18),
    cpr_to_18yrs = as.numeric(interval(CPR_DATE, WOB + years(18)), "years") %>% round(1),
#     study ends on 2021-05-31 when last CSC data was provided
# extra created to account for people documented in the last year
    cpr_to_stdy_end = as.numeric(interval(CPR_DATE, ymd("2021-05-31")), "years") %>% round(1),
    brth_to_cpr_yrs = as.numeric(as.duration(WOB %--% CPR_DATE), "years") %>% round(1),
#     events less than zero (i.e. -ve) are clipped to zero
    cpr_free_yrs = if_else(brth_to_cpr_yrs < 0, 0, brth_to_cpr_yrs),
  ) %>%
  glimpse

```

## Mother-Child cohort (cleaned after SB)

NB: Only individuals whose health records had ALF_STS_CD 1, 4 and 39 were extracted.

```{r}

# load & explore cleaned cohort data
d_cohort_clean <-
  qread("data/d_cohort_clean.qs") %>%
  # drop variables associated with CLA
  select(-starts_with("c_lac_")) %>% 
  glimpse

#   Our criteria for selecting samples from d_cohort_clean (for the survival analysis) is slightly different to SB's work. see their file 5b - select sample.r

fct_interval <- function(c_wob, c_gest_age, x_flg, x_date) {
    # for those who have x_flg == 1, which interval does x_date sit in:
    #
    #  -5 years                              pregnancy        week of
    # before wob                               start           birth
    #     |<------------------------------------>|<------------->|
    #

    z_lag_5y     <- as.Date(c_wob - dyears(5))
    z_preg_start <- c_wob - dweeks(c_gest_age)

    res <- case_when(
        x_flg == "No" ~ "never",
        x_flg == "Yes" & z_lag_5y     <= x_date & x_date < z_preg_start ~ "pre-pregnancy",
        x_flg == "Yes" & z_preg_start <= x_date & x_date <= c_wob       ~ "pregnancy",
        .default = as.character(x_flg)
    )

    res <- factor(res, c("never", "pre-pregnancy", "pregnancy"))
    return(res)
}

# check available variables
d_cohort_clean %>% glimpse

# study end date is 2021-05-31

# study population (roughly 1yr b4 to 1yr after the 2011 census)
surv_cohrt_ehr <-
  d_cohort_clean %>% #glimpse
#   c_born_in_study_window,
  filter(c_wob %within% (ymd("2010-01-01") %--% ymd("2012-12-31"))) %>%
  mutate(
#     child year of birth
    c_brth_yr_cat = year(c_wob) %>% factor,
#     mother's age at child birth
    m_age_at_birth = as.duration(m_wob %--% c_wob) %>% as.numeric(., "years") %>% round(1),
    m_age_at_birth_cat = case_when(m_age_at_birth < 16 ~ "Under16",
                                    m_age_at_birth >= 16 & m_age_at_birth < 18  ~ "16-17",
                                   m_age_at_birth >= 18 & m_age_at_birth < 20 ~ "18-19",
                                   m_age_at_birth >= 20 & m_age_at_birth < 25 ~ "20-24",
                                   m_age_at_birth >= 25 & m_age_at_birth < 30 ~ "25-29",
                                   m_age_at_birth >= 30 ~ "30plus",
                                   ) %>% fct_rev(),
    c_birth_weight_cat = case_when(c_birth_weight < 2500 ~ "Low",
                                    2500 <= c_birth_weight & c_birth_weight < 4000 ~ "Normal",
                                    c_birth_weight >= 4000 ~ "Heavy",
                                    .default = as.character(c_birth_weight)
                                    ) %>% factor %>%
  fct_relevel("Heavy", after = 2) %>% fct_relevel("Normal"),
    # categorise perinatal health (behaviour) as non, or pre/post-pregnancy
    m_alc_cat  = fct_interval(c_wob, c_gest_age, m_alc_flg,  m_alc_date),
    m_htn_cat  = fct_interval(c_wob, c_gest_age, m_htn_flg,  m_htn_date),
    m_mnt_cat  = fct_interval(c_wob, c_gest_age, m_mnt_flg,  m_mnt_date),
    m_sbm_cat  = fct_interval(c_wob, c_gest_age, m_sbm_flg,  m_sbm_date),
    
    # harmonise smoking measure, as we have several data sources and different
    # values correspond to being an ex-smoker or non-smoker
    m_smoking_cat = na_if(m_smoking_cat, ""),
    m_smk2_cat  = case_when(
            not_na(m_smoking_cat) ~ as.character(m_smoking_cat),
            m_smk_cat == "S"      ~ "Smk",
            m_smk_cat == "E"      ~ "Ex",
            m_smk_cat == "N"      ~ "Non",
            TRUE ~ "Unknown"
        ) %>%
        factor(c("Non", "Ex", "Smk", "Unknown")),
#     convert wimd from decile to quintile
      m_wimd2014_quintile = m_wimd2014_decile %>% fct_collapse(
            "1 - Most deprived"  = c("1 - Most deprived", "2"),
            "2"                  = c("3", "4"),
            "3"                  = c("5", "6"),
            "4"                  = c("7", "8"),
            "5 - Least deprived" = c("9", "10 - Least deprived")
        ) %>% fct_rev(),
# cuases singularity problems in the survival analysis
# %>% fct_na_value_to_level(level = "Missing"),
      # m_townsend2011_quintile = m_townsend2011_quintile %>%
  # fct_na_value_to_level(level = "Missing"),
#     will use more reliable data on ethnicity from the census
    c_ethn_cat =
            c_ethnicity_nm %>%
            as.character() %>%
            str_to_lower() %>%
            str_replace(" - ethnic category 2001 census", "") %>%
            str_replace(" - ethn categ 2001 census", "") %>%
            str_replace(".*black .+ white.*", "black and white") %>%
            str_replace(".*white .+ black.*", "white and black") %>%
            str_trim() %>%
            factor() %>%
            fct_collapse(
                "Asian" = c(
                    "any other asian background",
                    "bangladeshi",
                    "bangladeshi (asian or british)",
                    "bangladeshi or british bangladeshi",
                    "chinese",
                    "indian",
                    "indian (asian or british)",
                    "indian or british indian",
                    "pakistani",
                    "pakistani (asian or british)",
                    "pakistani or british pakistani",
                    "other asian background",
                    "other asian ethnic group"
                ),
                "Black" = c(
                    "any other black background",
                    "african",
                    "african(black or black british)",
                    "black african",
                    "black british",
                    "caribbean",
                    "caribbean (black or black british)",
                    "other (black or black british)",
                    "other black background"
                ),
                "Mixed" = c(
                    "any other mixed background",
                    "black and white",
                    "chinese and white",
                    "mixed white & asian",
                    "other ethnic, asian/white orig",
                    "other ethnic, mixed origin",
                    "other ethnic, other mixed orig",
                    "other mixed background",
                    "other mixed white",
                    "white and asian",
                    "white and black"
                ),
                "White" = c(
                    "any white background",
                    "other white background",
                    "other white british ethnic group",
                    "polish",
                    "white",
                    "white british",
                    "white irish",
                    "white - other",
                    "welsh"
                ),
                "Other" = c(
                    "any other ethnic group",
                    "any other group",
                    "other",
                    "other ethnic group"
                )
            ) %>%
            factor(levels = c(
                "White",
                "Asian",
                "Black",
                "Mixed",
                "Other"
            )) %>%
            fct_na_value_to_level("Unknown"),
        m_hh_child_n = tidyr::replace_na(m_hh_child_n, 0),
        m_hh_child_cat = case_when(
                m_hh_adult_n >= 1 & m_hh_child_n == 0 ~ "0 children",
                m_hh_adult_n >= 1 & m_hh_child_n >= 1 ~ "1+ children"
            ) %>%
            factor(), #%>%
        m_hh_adult_cat = case_when(
                m_hh_adult_n == 1 ~ "1 adult",
                m_hh_adult_n == 2 ~ "2 adults",
                m_hh_adult_n >= 3 ~ "3+ adults"
            ) %>%
            factor(), #%>%
          # fct_na_value_to_level("unknown"),
        m_hh_cat = case_when(
                m_hh_adult_n == 1 & m_hh_child_n == 0 ~ "1 adult only",
                m_hh_adult_n == 1 & m_hh_child_n >= 1 ~ "1 adult, 1+ children",
                m_hh_adult_n == 2 & m_hh_child_n == 0 ~ "2 adults only",
                m_hh_adult_n == 2 & m_hh_child_n >= 1 ~ "2 adults, 1+ children",
                m_hh_adult_n >= 3 & m_hh_child_n == 0 ~ "3+ adults only",
                m_hh_adult_n >= 3 & m_hh_child_n >= 1 ~ "3+ adults, 1+ children"
            ) %>%
            factor() %>%
          fct_na_value_to_level("unknown"),
      across(where(is.character), ~str_squish(.x) %>% fct_na_value_to_level(level = "None")
      )
  ) %>%
    # glimpse
  # frq(m_alc_cat, m_htn_cat, m_mnt_cat, m_sbm_cat #c_birth_weight_cat, c_ethn_cat
  #     )
  filter(
        c_stillbirth_flg == "No" & welsh_birth_flg == "Yes",
        not_na(c_wob) & not_na(c_gest_age) & not_na(c_preterm_birth_flg) & not_na(c_birth_weight), #c has birth info
        c_gp_flg == "Yes", #c_has_gp,
        has_m == "Yes",
        m_gp_flg == "Yes",#m_has_gp,
        m_age_at_birth >= 16,
    ) %>% 
  droplevels() %>%
#   select and rename pertinent variables
  select(
    # id
        c_alf_pe,
        c_alf_sts_cd,
    # outcome
        c_wob,
        c_death_date,
    # covariates
        c_brth_yr_cat,
        c_sex,
        c_ethn     = c_ethn_cat,
        c_cgm      = c_cgm_flg,
        c_ptb      = c_preterm_birth_flg,
        c_bw       = c_birth_weight_cat,
    m_alf_pe,
    m_alf_sts_cd,
        m_age      = m_age_at_birth,
        m_age_cat  = m_age_at_birth_cat,
        m_lrn      = m_lrn_flg,
        m_smk      = m_smk2_cat,
        m_hh_child = m_hh_child_cat,
        m_hh_adult = m_hh_adult_cat,
        m_hh       = m_hh_cat,
        m_alc      = m_alc_cat,
        m_mnt      = m_mnt_cat,
        m_htn      = m_htn_cat,
        m_sbm      = m_sbm_cat,
        m_twns2011 = m_townsend2011_quintile
    ) %>%
  glimpse

gc()

save.image()
```

## Join CSC data

```{r}

# study ends when an individual attains age 16, at which time they are no longer subject to CSC records

chld_cpr_dta <-
cinw_crcs_cpr %>% 
  # select(ALF_PE, CHILD_PROTECTION_REGISTER, DATE_OF_DATA_COLLECTION, WOB, GNDR_CD, drtn_yrs_cln) %>% 
  select(ALF_PE, CHILD_PROTECTION_REGISTER, CPR_DATE, cpr_free_yrs, cpr_to_stdy_end) %>%
  rename_with(~paste0("c_", .x)) %>% 
  glimpse

# 8296/22505 (36.9%) of CPR data has missing alf_pe
chld_cpr_dta %>% 
  ExpData(2)
# THis means about 8296 who may have CPR cannot be determined in the data

# study end date is 2021-05-31

ehr_cpr_cohrt <- 
surv_cohrt_ehr %>% 
  left_join(chld_cpr_dta, by = c("c_alf_pe" = "c_ALF_PE")) %>%
  # filter(!is.na(c_CHILD_PROTECTION_REGISTER)) %>% 
   mutate(
     c_CHILD_PROTECTION_REGISTER = fct_na_value_to_level(c_CHILD_PROTECTION_REGISTER, level = "False"),
     cpr_evnt = case_match(c_CHILD_PROTECTION_REGISTER,
                           "True" ~ 1,
                           "False" ~ 0),
     frst_evnt_cat = case_when(
       c_CHILD_PROTECTION_REGISTER == "False" & !is.na(c_death_date) ~ "Death",
       c_CHILD_PROTECTION_REGISTER == "True" ~ "CPR",
       c_death_date > c_CPR_DATE & c_death_date < ymd("2021-05-31") ~ "Death",
       .default = "Study End"
     ) %>% factor,
         # stdy_end_age_yrs = as.numeric(as.duration(c_wob %--% (c_wob+dyears(18))), "years") %>% round(1),
     dth_age_yrs = as.numeric(as.duration(c_wob %--% c_death_date), "years") %>% round(1),
     #     study ends on 2021-05-31 when last CSC data was provided
  # extra created to account for people documented in the last year
    stdy_end_age_yrs = as.numeric(as.duration(c_wob %--% ymd("2021-05-31")), "years") %>% round(1),
# #   if event doesn't happen before study end, individual survived throughout the study duration/followup
    c_cpr_2_stdy_end_yrs = tidyr::replace_na(c_cpr_to_stdy_end, 0),
    age_1st_evnt = case_when(
      frst_evnt_cat == "CPR" ~ `c_cpr_free_yrs`,
      frst_evnt_cat == "Death" & dth_age_yrs < stdy_end_age_yrs ~ `dth_age_yrs`,
      frst_evnt_cat == "Study End" | dth_age_yrs >= stdy_end_age_yrs ~ `stdy_end_age_yrs`,
    )
     ) %>% 
  # some individuals have more than one date for first CPR registration, so dedup
  slice_min(order_by = c_CPR_DATE, n = 1, by=c_alf_pe) %>%
  glimpse

```

The (exponentiated) regression coefficients from cox regression are interpreted as
hazard-ratios.

# Cohort updated with census records

Census variables are more reliable, so pertinent ones in SB data will be swapped with those from the census. This includes the main exposure

## Census-updated cohort table

Swap ethnicity, m_hh_child, m_hh_adult, m_twns2011

The 4 Dimensions considered for HH deprivation (DEPRIVED) are: 1, Education:
DEPEDHUK11 [Household education deprivation], 2, Employment: DEPEMHUK11 [Employment
deprivation], 3, Health and Disability: DEPHDHUK11 [Health and Disability
Deprivation], 4, Housing: DEPHSHUK11 [Housing deprivation]

The main exposure One option is to use number of dimensions of HH deprivation
(DEPRIVED).

Another is to use an intersection of all the dimensions of deprivation as the main
exposure, so that the study can speak to both particular dimensions and their
combined (or interaction) effects.

Replace: 
- m_hh_adults with ADTHUK11 [Number of adults in household], grouped as 1,
2, and 3+ 
- ethnicity with AETHPUK11 [Alternative Ethnicity Ethnic group
(Alternative)]: Arab, 'Caribbean, African, Black British and other Black', Chinese,
Indian, Mixed/multiple ethnic groups, No code required, Other ethnic group, Pakistani and other Asian/Asian British, White. "No code required" recoded as "Unknown".
m_hh_adult with DPCFAMUK11 [Dependent children in family]: 0, 1, 2, 3+. Or count
using DCHPUK11 [Dependent child indicator].

Add: 
- CFSPUK11 [Child family status]: Cohabiting couple family, Lone parent family,
Married couple family, Same-sex civil partnership family, 
- MARSTAT [Marital and civil partnership status]: Divorced Formerly in a same-sex civil partnership which is now legally dissolved In a registered same-sex civil partnership Married Separated (but still legally in a same-sex civil partnership) Separated (but still legally married) Single (never married or never registered a same-sex civil partnership) Surviving partner from a same-sex civil partnership Widowed

Categories in MARSTAT have been recoded to enhance clarity, Where suffix is
"Same-sex", this applies to "same-sex civil partnerships"

This might be more relevant than the former because unlike the former, it provides
marital information whether the individual has a child or not. However, there is a
need to recode categories into fewer groups.

-   Optionally: HLQPUK11 [Highest level of qualification]: Level 4, Level 3,
    Apprenticeship, Level 2, Level 1, Others, No qualifications, Not applicable. But
    this might collide with HH deprivation count/dimensions, but this is the case
    with aspects like health (used in SB's study).
-   Optionally: LRESPUK11 [Length of residence in the UK]: UK-born, <=2 yrs, 2 -
    <5yrs, 5 - <10yrs, >=10yrs

```{r}
# A near-consistent amount of missingness across these Census categories reflects a systematic limitation of the available linked source data. Missing values are either <NA> | 837234 | 56.08 | or | 836185 | 56.01 |    <NA> | or | 837079 | 56.07 |    <NA> |,

# Missing values are retained as NA because recoding them to a new category would introduce an artificial category.

mother_ses_cens11 <- # n = 96,442
mother_child_cens11 %>% #n = 1,492,953
  semi_join(ehr_cpr_cohrt, by = c("ALF_PE" = "m_alf_pe")) %>% 
    # glimpse
  select(ALF_PE, ADTHUK11, AETHPUK11, DPCFAMUK11, CFSPUK11, MARSTAT, DEPRIVED, DEPEDHUK11, DEPEMHUK11, DEPHDHUK11, DEPHSHUK11, LRESPUK11, m_edu = HLQPUK11) %>% #replace CFSPUK11 with MARSTAT
  # frq()
  mutate(
    m_edu = fct_recode(m_edu,
                       "None" = "No qualifications",
                       "None" = "Not applicable"
                       ), #%>% 
    m_ethnic = AETHPUK11 %>% 
#       'No code require' no recoded to NA instead of 'Unknown' to mitigate problems of colinearity and sparse levels
      fct_recode(NULL = "No code required") %>%
      fct_infreq(), #%>% 
    m_hh_adult = case_match(ADTHUK11,
                            c(0,1) ~ "0/1", #grouped because of small numbers of the 0 category
                               c(3:50) ~ "3+",
                               .default = as.character(ADTHUK11)
    ) %>% factor %>%
      fct_relevel("2"), #makes this the reference level
#     'No code required' changed to 0
    m_hh_child_depndt = case_when(
      grepl("^No dependent", DPCFAMUK11) ~ "0",
      grepl("^One dependent", DPCFAMUK11) ~ "1",
      grepl("^Two dependent", DPCFAMUK11) ~ "2",
      grepl("^Three or more dependent", DPCFAMUK11) ~ "3+",
      grepl("^No code", DPCFAMUK11) ~ "0", #otherwise 99
      .default = DPCFAMUK11
    ) %>% factor, #%>%
    m_hh_deprv_no = case_when(
      grepl(".not deprived.", DEPRIVED) ~ "0",
      grepl("deprived in 1 dimension$", DEPRIVED) ~ "1",
      grepl("deprived in 2 dimensions$", DEPRIVED) ~ "2",
      grepl("deprived in 3 dimensions$", DEPRIVED) ~ "3",
      grepl("deprived in 4 dimensions$", DEPRIVED) ~ "4",
      grepl("^No code", DEPRIVED) ~ "0", #otherwise 99
      .default = DEPRIVED
    ) %>% factor, #%>%
     hh_edu_dprvd = case_when(
       grepl("Yes", DEPEDHUK11) ~ "education",
       grepl("^No code", DEPEDHUK11) ~ NA, #was no
       ),# %>% fct_na_value_to_level("No"),
      hh_emp_dprvd = case_when(
       grepl("Yes", DEPEMHUK11) ~ "employment",
       grepl("^No code", DEPEMHUK11) ~ NA, #was no
       ),# %>% fct_na_value_to_level("No"),
      hh_hlt_dprvd = case_when(
       grepl("Yes", DEPHDHUK11) ~ "health",
       grepl("^No code", DEPHDHUK11) ~ NA, #was no
       ),# %>% fct_na_value_to_level("No"),
      hh_hous_dprvd = case_when(
       grepl("Yes", DEPHSHUK11) ~ "housing",
       grepl("^No code", DEPHSHUK11) ~ NA, #was no
       ),# %>% fct_na_value_to_level("No"),
    m_fam_status = CFSPUK11 %>%
  fct_relabel(~gsub("^Has children: ", "", .x)) %>%
  fct_relabel(~gsub(" family$", "", .x)) %>% 
      fct_recode(NULL = "No code required") %>% 
  fct_relevel("Married couple") %>% 
      fct_relevel("Does not have children", after = 4),
       # fct_na_value_to_level("Unknown"),
    m_mar_status = fct_recode(MARSTAT,
      "Divorced (same-sex)" = "Formerly in a same-sex civil partnership which is now legally dissolved",
     "Married (Same-sex)" = "In a registered same-sex civil partnership",
     "Separated (same-sex)" = "Separated (but still legally in a same-sex civil partnership)",
     "Separated" = "Separated (but still legally married)",
     "Single" = "Single (never married or never registered a same-sex civil partnership)",
     "Widowed (same-sex)" = "Surviving partner from a same-sex civil partnership",
    ) %>% 
  fct_infreq() %>% 
      # fct_relevel("Unknown", after = Inf) %>% 
  fct_relevel("Widowed", after = 4) %>% 
  fct_relevel("Married"), #the reference category
    m_uk_stay_len = fct_recode(
      LRESPUK11,
      "UK-born" = "Born in the UK",
      "10+ yrs" = "Resident in UK: 10 years or more",
      "5 - <10 yrs" = "Resident in UK: 5 years or more but less than 10 years",
      "<5 yrs" = "Resident in UK: 2 years or more but less than 5 years",
      "<5 yrs" = "Resident in UK: Less than 2 years",
      NULL = "No code required"
    ) %>% 
      fct_relevel(c("UK-born", "10+ yrs", "5 - <10 yrs", "<5 yrs")),
  ) %>% 
  mutate(
    m_hh_deprv_dms = paste0(hh_edu_dprvd, "+", hh_emp_dprvd, "+", hh_hlt_dprvd, "+", hh_hous_dprvd) %>%
      str_replace_all("\\+NA|NA\\+", "") %>%
      case_match(., "NA"~"None", .default = .) %>%
      factor %>%
      fct_infreq() %>% 
      fct_relevel("housing", after = 4) %>%
      fct_relevel("education+employment+health", "employment+health+housing", after = 11) %>%
      fct_relevel("education+employment+health+housing", after = Inf),
  ) %>% 
#   drop processed variables
  select(-c(ADTHUK11, AETHPUK11, DPCFAMUK11, DEPRIVED, DEPEDHUK11, DEPEMHUK11, DEPHDHUK11, DEPHSHUK11, MARSTAT, LRESPUK11, CFSPUK11, hh_edu_dprvd, hh_emp_dprvd, hh_hlt_dprvd, hh_hous_dprvd)) %>% 
glimpse

ehr_cpr_cens11_cohrt <- # n=106,365
ehr_cpr_cohrt %>% #glimpse()
  # drop variables being replaced with census ones
  select(-c(m_hh_child, m_hh_adult, m_hh)) %>% 
  left_join(mother_ses_cens11, by = c("m_alf_pe" = "ALF_PE")) %>% 
  mutate(
     frst_cpr = frst_evnt_cat=="CPR",
     m_twns2011 = fct_rev(m_twns2011)
  ) %>%
#   variable selection with renaming
  select(-c(c_cpr_to_stdy_end, c_cpr_2_stdy_end_yrs), c_cpr_evnt = cpr_evnt, c_frst_evnt_cat=frst_evnt_cat, c_dth_age_yrs=dth_age_yrs, c_stdy_end_age_yrs=stdy_end_age_yrs, c_age_1st_evnt = age_1st_evnt, c_frst_cpr=frst_cpr) %>%
  filter(!is.na(c_sex)) %>% #removed due to small numbers (n = 4)
  droplevels() %>%
  glimpse

# for renaming variables
var_list_full <- list(
# dummy
    # "dummy"              = "Total",
    # child baseline measures
    "c_brth_yr_cat" = "Child birth year",
    "c_sex"              = "Child sex",
    "c_ethn"             = "Child ethnicity",
    "c_ptb"              = "Child preterm birth",
    "c_bw"               = "Child birth weight",
    "c_cgm"              = "Child congenital anomalies",
    "c_CHILD_PROTECTION_REGISTER" = "Child CPR registration",
    "c_frst_evnt_cat" = "Child 1st event",
    "c_age_1st_evnt" = "Child median age at CPR registration",
    "m_age_cat"          = "Mother age at birth (category)",
    "m_lrn"              = "Mother learning difficulties/disabilities",
    "m_smk"              = "Mother smoking",
    "m_alc"              = "Mother alcohol problems",
    "m_mnt"              = "Mother mental health issues",
    "m_htn"              = "Mother hypertension",
    "m_sbm"              = "Mother substance use",
    "m_ethnic"           = "Mother ethnicity",
    "m_edu"             = "Mother education",
    "m_hh_child_depndt"  = "Mother HH dependent children",
    "m_hh_adult"         = "Mother HH adults",
    "m_mar_status"       = "Mother marital status",
    "m_fam_status" = "Mother family status",
    "m_uk_stay_len"      = "Mother length of UK stay",
    "m_hh_deprv_no"      = "Mother HH deprivation count",
    "m_hh_deprv_dms"     = "Mother HH deprivation dimensions",
    # "m_wimd2014"         = "Mother WIMD 2014 quintile",
    "m_twns2011"         = "Mother Townsend 2011 quintile"
)

names(var_list_full)

tic()
print(Sys.time())
ehr_cpr_cens11_cohrt %>% 
  select(all_of(names(var_list_full))) %>% 
  # glimpse
  labelled::set_variable_labels(
    .labels = var_list_full,
    .strict = F
  ) %>% 
  # glimpse
  tbl_summary(
    # by = mother_child,
    type = list(where(is.logical) ~ "categorical",
                all_dichotomous() ~ "categorical"),
    missing_text = "NA (Missing)",
    missing_stat = "{N_miss} ({p_miss}%)",
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = everything() ~ c(0,2)
  ) %>%
  bold_labels() %>% 
  modify_caption("**Characteristics of Birth Cohort A: Born between 2010 and 2012 followed until Study End (31 March, 2021)**") %>% 
  as_gt() %>% #convert output to a gt table for onward saving.
  gt::tab_header(title = gt::md("Characteristics of the birth (prospective) cohort for this study")) %>%
  gt::gtsave("outputs\\cpr_birth_cohort_2010_2012.docx") #save produced table as word
print(Sys.time())
toc()


gc()
save.image()

```

Future extension of study can consider: OCCPUK113 [Occupation (Major group)] &
NSSEC [National Statistics Socio-economic Classification]
# Data exploration
## Correlation analyses
```{r}
# function to rename discrete x|y axis using a named list
# not needed as can be done with map_chr, such as map_chr(var_list_full, identity)
# make_axis_labels <- 
#   function(label_lookup){
#     function(breaks){
#       vapply(breaks,
#              function(x){
#                if(x %in% names(label_lookup)){
#                  as.character(label_lookup[[x]])
#                } else{
#                  x #retain lables not present in the lookup list
#                }
#              },
#              character(1))
#     }
#   }
# for renaming variables v2
var_list_full2 <- list(
    # child baseline measures
    "c_brth_yr_cat" = "Child birth year",
    "c_sex"              = "Child sex",
    "c_ethn"             = "Child ethnicity",
    "c_ptb"              = "Child preterm birth",
    "c_bw"               = "Child birth weight",
    "c_cgm"              = "Child congenital anomalies",
    "c_CHILD_PROTECTION_REGISTER" = "CPR registration",
    "c_frst_evnt_cat" = "Child 1st event",
    "m_age_cat"          = "Mother age at birth",
    "m_lrn"              = "Mother learning\ndifficulties/disabilities",
    "m_smk"              = "Mother smoking",
    "m_alc"              = "Mother alcohol problems",
    "m_mnt"              = "Mother mental health\nissues",
    "m_htn"              = "Mother hypertension",
    "m_sbm"              = "Mother substance use",
    "m_ethnic"           = "Mother ethnicity",
    "m_edu"              = "Mother education",
    "m_hh_child_depndt"  = "Mother HH dependent\nchildren",
    "m_hh_adult"         = "Mother HH adults",
    "m_mar_status"       = "Mother marital status",
    "m_fam_status"       = "Mother family status",
    "m_uk_stay_len"      = "Mother length of UK stay",
    "m_hh_deprv_no"      = "Mother HH deprivation\ncount",
    "m_hh_deprv_dms"     = "Mother HH deprivation\ndimensions",
    # "m_wimd2014"         = "Mother WIMD 2014 quintile",
    "m_twns2011"         = "Mother Townsend 2011 quintile"
)

names(var_list_full2)

# full variables all
vars_all <- c(
  "c_CHILD_PROTECTION_REGISTER"  ,
  "c_sex",
    "c_ethn",
    "m_age_cat",
    "m_edu",
    "m_hh_child_depndt",
    "m_hh_adult",
    "m_mar_status", #**
    "m_hh_deprv_dms",
    "m_hh_deprv_no", #**
    "c_cgm",
    "c_ptb",
    "c_bw",
    "m_alc",
    "m_lrn",
    "m_mnt",
    "m_sbm",
    "m_smk"
    )

# map 'categorical' to each column name
clmn_types <- set_names(
  rep(list("categorical"), ncol(ehr_cpr_cens11_cohrt)),
  names(ehr_cpr_cens11_cohrt)
)

# appears like a very time consuming operation (over one hour)
tic()
Sys.time()
cars_corr_all <- 
ehr_cpr_cens11_cohrt %>%
  select(all_of(vars_all)) %>%
  smart_cormat(
    types = clmn_types,
    assume_latent_normal = F,
               verbose = T
               )
Sys.time()
toc() #3.03 sec elapsed

# Tidy and export correlation matrix to csv file
cars_corr_all |>
  tidy() |>
  mutate(var_x = recode_values(var_x, from = names(var_list_full2),
                       to = var_list_full2),
         var_y = recode_values(var_y, from = names(var_list_full2),
                       to = var_list_full2)
         ) |> #glimpse()
  write_csv("outputs\\cars_corr_all_plt.csv")
  # view()

# create plot
cars_corr_all_plt <- 
  ggcor_heatmap(cars_corr_all,
                lower_only = T,
                text_size = 4) +
  scale_x_discrete(labels = map_chr(var_list_full2, identity)) +
  scale_y_discrete(labels = map_chr(var_list_full2, identity)) +
  # Replace legend
  scale_fill_gradient2(limits = c(0, 1),
                       breaks = seq(0, 1, by = 0.2),
                       high = muted("red"),
                       na.value = NA,
                       name = "Cramer's V"
                       ) +
  # labs(
  #   title = "Correlation Matrix"
  # ) +
  theme(axis.text = element_text(size = 12),
        legend.position = "right",
        legend.justification.right = "bottom",
        legend.key.height = grid::unit(2, "cm")
        )

ggsave(
  filename = "outputs\\cars_corr_all_plt.png",
  plot = cars_corr_all_plt,
  # scale = 0.8,
  height = 30,
  width = 35,
  units = 'cm',
  dpi = 900
)

gc()
save.image()

# cars_corr_all |>
#   tidy() |> 
#   select(-c("method", "rationale", "p.value", "p_method", "null_hypothesis", "p_interpretation")) |> 
#   glimpse()

```

# Survival plot: Smoothening helper function
```{r}
# SDC-friendly smoothed survival/risk plot helper.
# The survival model is unchanged; smoothing/noise is applied only to the plotted curve.

plot_sdc_smooth_survfit <- function(survfit_obj,
                                    type = c("risk", "survival"),
                                    method = "loess",
                                    formula = y ~ x,
                                    span = 0.10,
                                    smooth_n = 100,
                                    method.args = list(),
                                    add_noise = FALSE,
                                    noise_scope = c("tail", "all"),
                                    noise_tail_n = 5,
                                    noise_sd = 0.0005,
                                    noise_seed = NULL,
                                    show_ci = TRUE,
                                    subtitle = NULL,
                                    caption = NULL) {
  type <- match.arg(type)
  noise_scope <- match.arg(noise_scope)

  plot_data <-
    ggsurvfit::tidy_survfit(survfit_obj) #%>%
  
  if ("strata" %in% names(plot_data)) {
    plot_data <- plot_data |> dplyr::mutate(strata = as.character(strata))
  }else{
    plot_data <- plot_data |> dplyr::mutate(strata = "Overall")
  }

  if (!"conf.low" %in% names(plot_data)) {
    plot_data <- plot_data %>% dplyr::mutate(conf.low = NA_real_)
  }

  if (!"conf.high" %in% names(plot_data)) {
    plot_data <- plot_data %>% dplyr::mutate(conf.high = NA_real_)
  }

  # Convert survival probability to cumulative risk when requested.
  if (type == "risk") {
    plot_data <-
      plot_data %>%
      dplyr::mutate(
        conf.low_surv = .data$conf.low,
        conf.high_surv = .data$conf.high,
        estimate = 1 - .data$estimate,
        conf.low = 1 - .data$conf.high_surv,
        conf.high = 1 - .data$conf.low_surv
      )
  }

  # Optional SDC noise: usually apply only to the final time points, where
  # small numbers at risk can make exact curve steps more disclosure-sensitive.
  if (add_noise) {
    if (!is.null(noise_seed)) {
      set.seed(noise_seed)
    }

    plot_data <-
      plot_data %>%
      dplyr::group_by(.data$strata) %>%
      dplyr::arrange(.data$time, .by_group = TRUE) %>%
      dplyr::mutate(
        sdc_noise_zone = dplyr::case_when(
          noise_scope == "all" ~ TRUE,
          dplyr::row_number() > dplyr::n() - noise_tail_n ~ TRUE,
          TRUE ~ FALSE
        ),
        sdc_noise = dplyr::if_else(
          .data$sdc_noise_zone,
          stats::rnorm(dplyr::n(), mean = 0, sd = noise_sd),
          0
        ),
        dplyr::across(
          c("estimate", "conf.low", "conf.high"),
          ~ dplyr::if_else(
            .data$sdc_noise_zone & !is.na(.x),
            pmin(pmax(.x + .data$sdc_noise, 0), 1),
            .x
          )
        ),
        conf.low = pmin(.data$conf.low, .data$estimate, na.rm = TRUE),
        conf.high = pmax(.data$conf.high, .data$estimate, na.rm = TRUE)
      ) %>%
      dplyr::ungroup()
  }

  p <-
    ggplot2::ggplot(
      plot_data,
      ggplot2::aes(
        x = .data$time,
        y = .data$estimate,
        colour = .data$strata,
        fill = .data$strata,
        group = .data$strata
      )
    )

  if (show_ci) {
    p <-
      p +
      ggplot2::geom_ribbon(
        ggplot2::aes(ymin = .data$conf.low, ymax = .data$conf.high),
        alpha = 0.18,
        colour = NA
      )
  }
# scales::
  # scale_x
  p +
    ggplot2::geom_smooth(
      method = method,
      formula = formula,
      span = span,
      n = smooth_n,
      method.args = method.args,
      se = FALSE,
      linewidth = 0.5,
      na.rm = TRUE
    ) +
    ggplot2::scale_y_continuous(labels = scales::label_percent()
                                ) +
    # modify with new time span
    # ggplot2::scale_x_continuous(limits = c(0,11), breaks = seq.int(0,11,1)) +
    ggplot2::labs(
      subtitle = subtitle,
      caption = caption,
      y = ifelse(type == "risk", "% Risk (CPR Registration)", "Survival Probability"),
      x = "Follow-up Age (years)",
      colour = NULL,
      fill = NULL
    ) +
    ggplot2::theme_bw() +
    ggplot2::theme(legend.position = "bottom")
}

#----------------------------------------------
# Replace the unadjusted plot block with:
# all_klpn_plts[[paste0(covar_all[i])]] <-
#   plot_sdc_smooth_survfit(
#     survfit_obj = all_klpn_mdls[[i]],
#     type = "risk",
#     method = "loess",
#     span = 0.75,
#     add_noise = FALSE,
#     subtitle = paste0(names(covar_all[i]))
#   )

# If you want the optional final-tail SDC jitter:
# all_klpn_plts[[paste0(covar_all[i])]] <-
#   plot_sdc_smooth_survfit(
#     survfit_obj = all_klpn_mdls[[i]],
#     type = "risk",
#     method = "loess",
#     span = 0.75,
#     add_noise = TRUE,
#     noise_scope = "tail",
#     noise_tail_n = 4,
#     noise_sd = 0.0005,
#     noise_seed = 1409,
#     subtitle = paste0(names(covar_all[i]))
#   )

# For adjusted plots, replace only the ggsurvfit(...) + ... plotting section with:
# cxph_adj_demo_plts[[adj_mdl_nme]] <-
#   plot_sdc_smooth_survfit(
#     survfit_obj = cxph_adj_demo_mdls[[adj_mdl_nme]],
#     type = "risk",
#     method = "loess",
#     span = 0.75,
#     add_noise = FALSE,
#     subtitle = strata_nme,
#     caption = paste0("Demography Adjusted Model (", mdl_nme, ")")
#   )
# 
# cxph_adj_full_plts[[adj_mdl_nme]] <-
#   plot_sdc_smooth_survfit(
#     survfit_obj = cxph_adj_full_mdls[[adj_mdl_nme]],
#     type = "risk",
#     method = "loess",
#     span = 0.75,
#     add_noise = FALSE,
#     subtitle = strata_nme,
#     caption = paste0("Fully Adjusted Model (", mdl_nme, ")")
#   )

# For stronger smoothing, try span = 0.9. For GAM smoothing instead of LOESS:
# plot_sdc_smooth_survfit(
#   survfit_obj = all_klpn_mdls[[i]],
#   type = "risk",
#   method = "gam",
#   formula = y ~ s(x, bs = "cs"),
#   method.args = list(method = "REML"),
#   subtitle = paste0(names(covar_all[i]))
# )
```

# Kaplan-Meier Plots

## Risk (%)
Risk is plotted because it is easy to zoom into its y values
```{r}

# probability of being on the CPR at each age
## All potential exposures

covar_all[[14]]
covar_all |> names()

for(i in seq_along(covar_all)) {

  if (i == 1) {
    print(Sys.time())
    tictoc::tic()
    
    #     No. of surv models
    tbl_cnt <- length(covar_all)
    
    all_klpn_mdls = list()
    all_klpn_tbls = list()
    all_klpn_plts = list()
    gt_suvfit_tbl = list()
      
    #make outputs directory (if it doesn't already exist)
    unadj_surv_plts <- "outputs\\rsk_plts_unadjs_sdc"
    
    if (!dir.exists(unadj_surv_plts)) {
      dir.create(unadj_surv_plts)
    }
    
  }

  kp_fmla = paste0("\nsurvival::Surv(c_age_1st_evnt, c_frst_cpr) ~ ", covar_all[[i]])
  cat("\nFormula", kp_fmla)
  cat("\n\tPlotting", covar_all[i], "...\n")
  
  #   create model
  all_klpn_mdls[[paste0(covar_all[i])]] <-
    ggsurvfit::survfit2(as.formula(kp_fmla),
                        data = ehr_cpr_cens11_cohrt)
  
#   publish survfit table
    # NOT NOW NEEDED DUE TO SAILS SDC 
  # gt_suvfit_tbl[[paste0(covar_all[i])]] <- 
  # gtsummary::tbl_survfit(
  #   all_klpn_mdls[[i]],
  #   times = seq.int(1,11,2),
  #   estimate_fun = label_style_percent(digits = 1),
  #   label = var_list_full,
  #   label_header = "**{time} Years**"
  # )
  
  #   tidy tables of the model, for flexible plotting elsewhere
  # NOT NOW NEEDED DUE TO SAILS SDC 
  # all_klpn_tbls[[paste0(covar_all[i])]] <-
  #   all_klpn_mdls[[i]] %>%
  #   tidy_survfit()
  
  #   create plots from models
  ## CURVES NOT SMOOTHED
  # all_klpn_plts[[paste0(covar_all[i])]] <-
  #   all_klpn_mdls[[i]] %>%
  #   #     risk plots can be replaced with  Cummulative hazard ("cumhaz") or survival probability ("survival") plots
  #   ggsurvfit::ggsurvfit(type = "risk") +
  #   add_confidence_interval() +
  #   stat_smooth(x = "time", y = "estimate") +
  #   # add_censor_mark(stroke = 0.3, alpha = 0.5) + #offed cos of SDC
  #   scale_ggsurvfit() +
  #   add_pvalue("annotation") +
  #   labs(subtitle = paste0(names(covar_all[i])),
  #        y = "% Risk (CPR Registration)",
  #        x = "Follow-up Age (years)")
  
  ## CURVES NOW SMOOTHED
  all_klpn_plts[[paste0(covar_all[i])]] <-
  plot_sdc_smooth_survfit(
    survfit_obj = all_klpn_mdls[[i]],
    type = "risk",
    method = "loess",
    span = 0.15, #reduce to reduce smoothing
    smooth_n = 25, #tends to affect the CIs
    add_noise = T,
    noise_scope = "tail",
    noise_tail_n = 3,
    noise_sd = 0.00005,
    show_ci = T,
    subtitle = paste0(names(covar_all[i]))
  ) +
    ggplot2::scale_x_continuous(#limits = c(0,11),
                                breaks = seq.int(0,10,2),
                                expand = expansion(add = 0.1)
                                )

  #   adjust plot's y coordinates of relevant exposures (by zooming)
  if (covar_all[[i]] %in% c("m_mar_status")) {
    all_klpn_plts[[i]] <-
      all_klpn_plts[[i]] + coord_cartesian(ylim = c(0, 0.175))
  }
  if (covar_all[[i]] == "m_fam_status") {
    all_klpn_plts[[i]] <-
      all_klpn_plts[[i]] + coord_cartesian(ylim = c(0, 0.05))#
  }
  if (covar_all[[i]] == "m_ethnic") {
    all_klpn_plts[[i]] <-
      all_klpn_plts[[i]] + coord_cartesian(ylim = c(0, 0.035))
  }
  if (covar_all[[i]] == "c_ethn") {
    all_klpn_plts[[i]] <-
      all_klpn_plts[[i]] + coord_cartesian(ylim = c(0, 0.027))
  }
  if (covar_all[[i]] == "m_hh_deprv_no") {
    all_klpn_plts[[i]] <-
      all_klpn_plts[[i]] + coord_cartesian(ylim = c(0, 0.1)) #8
  }
  if (covar_all[[i]] == "m_hh_deprv_dms") {
    all_klpn_plts[[i]] <-
      all_klpn_plts[[i]] + coord_cartesian(ylim = c(0, 0.1)) #8
  }
  if (covar_all[[i]] == "c_cgm") {
    all_klpn_plts[[i]] <-
      all_klpn_plts[[i]] + coord_cartesian(ylim = c(0, 0.03))
  }
  
  #   format output file name by removing spaces, this also ensures no new output directory is accidentally created
  fle_nme <- str_replace_all(names(covar_all[i]), "[ |\\/|\\\\]", "_")
  
  #   plots needing wider width
  if (covar_all[[i]] %in% c("m_fam_status", "m_mar_status")) {
    wdth = 20
  } else if (covar_all[[i]] %in% c("m_hh_deprv_dms", "m_ethnic")) {
    wdth = 25
  }
  else{
    wdth = 15
  }

  cat("\n\tSaving plot...")
  
  ggsave(
    filename = paste0(unadj_surv_plts, "\\", fle_nme, ".png"),
    plot = all_klpn_plts[[i]],
    scale = 0.9,
    height = 15,
    width = wdth,
    units = 'cm',
    dpi = 600
  )
  
  if (i == tbl_cnt) {
    # TIDY TABLES NO LONGER SAVED BECAUSE WILL NOT PASS SAIL'S SDC
#     cat("\nSaving tidy tables of all plots ...\n")
# 
# #     save publication-ready table as word doc
#     gtsummary::tbl_stack(gt_suvfit_tbl)%>% 
#   as_gt() %>% 
#   gt::tab_header(title = gt::md("CPR Survival estimates of the study population for each exposure (unadjusted)")) %>% 
#   ## convert output to a gt table for onward saving.
#   gt::gtsave(paste0(unadj_surv_plts,"\\kpln_mr_plt_estimates_all.docx"))
    
    print(Sys.time())
    print(tictoc::toc())
    
    cat("\nSaving workspace ...\n")
    rm(tbl_cnt, fct_no, fle_nme, kp_fmla)
    gc()
    # save.image()
  }
  
}

# all_klpn_tbls[[1]]

```


# CoxPH Regression

## Bivariate (unadjusted) models

This should come first because it inform what variables to stratiy the kaplan-meier
plots

```{r}
# ehr_cpr_cens11_cohrt %>% glimpse

# list all potantial exposures/covariates
covar_all <- c(
    "Child Sex" = "c_sex",
    "Child ethnicity"="c_ethn",
    "Child congenital anomalies"="c_cgm",
    "Child preterm birth"="c_ptb",
    "Child birth weight"="c_bw",
    "Mother age at birth (category)"="m_age_cat",
    "Mother ethnicity"="m_ethnic",
    "Mother eductation"="m_edu",
    "Mother HH dependent children"="m_hh_child_depndt",
    "Mother HH adults"="m_hh_adult",
    "Mother marital status"="m_mar_status",
    "Mother family status"="m_fam_status",
    "Mother length of UK stay"="m_uk_stay_len",
    "Mother HH deprivation count"="m_hh_deprv_no",
    "Mother HH deprivation dimensions"="m_hh_deprv_dms",
    "Mother Townsend 2011 quintile"="m_twns2011",
    "Mother alcohol problems"="m_alc",
    "Mother learning difficulties/disabilities"="m_lrn",
    "Mother mental health issues"="m_mnt",
    "Mother substance use"="m_sbm",
    "Mother smoking"="m_smk"
    )

# make formula
for(i in seq_along(covar_all)) {
  if (i == 1) {
    cx_mdl_ft_a <- list()
    mdl_params_a <- list()
  }
  
  #   prep formula
  frml_a <- as.formula(paste0("Surv(c_age_1st_evnt, c_frst_cpr) ~ ", covar_all[[i]]))
  
  cat("\nModel", i, "\n")
  print(frml_a)
  
  #     fit bivariate survival model
  cx_mdl_ft_a[[covar_all[[i]]]] <- survival::coxph(formula = frml_a,
                                      data    = ehr_cpr_cens11_cohrt,
                                      id      = c_alf_pe)
  
  cat("Processing model parameters...")
  #     obtain and process counts per group
  var_cat_cnt <-
    ehr_cpr_cens11_cohrt %>%
    select(covar_all[[i]]) %>%
    na.omit() %>% #because analysis is often for complete cases
    frq(get(covar_all[[i]])) %>%
    data.frame() %>%
    select(val, frq) %>%
    mutate(val = as.character(val))
  
  #     obtain and process model parameters
  mdl_params_a[[names(covar_all[i])]] <-
    model_parameters(
      cx_mdl_ft_a[[i]],
      #18, 9
      exponentiate = T,
      include_info = T,
      pretty_names = F,
      include_reference = T
    ) %>%
    mutate(Parameter = str_replace_all(Parameter, covar_all[[i]], ""),
           group = covar_all[[i]]) %>%
    left_join(var_cat_cnt, by = c("Parameter" = "val")) %>%
    relocate(group)
  
  if (i == length(covar_all)) {
    cat("Saving all model tables as csv...")
    
    mdl_params_a_df <- bind_rows(mdl_params_a, .id = "Exposure") %>%
      select(-c(z, df_error)) %>% 
      mutate(sig = case_when( p < 0.001 ~ "***",
                              p < 0.01 ~ "**",
                              p < 0.05 ~ "*",
                              p < 0.1 ~ "."
                              )
             ) %>% 
      write_csv(file = "outputs\\bivar_coxph_cpr_all.csv")
    print(mdl_params_a_df, n = 100)
    
#     clean workspace
    rm(var_cat_cnt)
  }
}

# obtain performance stats from the survival models
## tabulate the performance of multiple models
compare_performance(cx_mdl_ft_a) %>%
  write_csv(file = "outputs\\bivar_coxph_cpr_perf.csv")

# tabulate bivariate models using gtsummary

# for renaming variables
mdl_tab_lbl <- list(
# dummy
    # "dummy"              = "Total",
    # child baseline measures
    "c_brth_yr_cat" = "Child birth year",
    "c_sex"              = "Child sex",
    "c_ethn"             = "Child ethnicity",
    "c_ptb"              = "Child preterm birth",
    "c_bw"               = "Child birth weight",
    "c_cgm"              = "Child congenital anomalies",
    "c_CHILD_PROTECTION_REGISTER" = "Child CPR registration",
    "c_frst_evnt_cat" = "Child 1st event",
    "c_age_1st_evnt" = "Child median age at CPR registration",
    "m_age_cat"          = "Mother age at birth (category)",
    "m_lrn"              = "Mother learning difficulties/disabilities",
    "m_smk"              = "Mother smoking",
    "m_alc"              = "Mother alcohol problems",
    "m_mnt"              = "Mother mental health issues",
    "m_htn"              = "Mother hypertension",
    "m_sbm"              = "Mother substance use",
    "m_ethnic"           = "Mother ethnicity",
    "m_edu"             = "Mother eductation",
    "m_hh_child_depndt"  = "Mother HH dependent children",
    "m_hh_adult"         = "Mother HH adults",
    "m_mar_status"       = "Mother marital status",
    "m_fam_status" = "Mother family status",
    "m_uk_stay_len"      = "Mother length of UK stay",
    "m_hh_deprv_no"      = "Mother HH deprivation count",
    "m_hh_deprv_dms"     = "Mother HH deprivation dimensions",
    # "m_wimd2014"         = "Mother WIMD 2014 quintile",
    "m_twns2011"         = "Mother Townsend 2011 quintile"
)

# the following chunk is as after taught, otherwise could have been merged with the main loop above.
for(i in seq_along(cx_mdl_ft_a)) {
  if (i == 1) {
    bivar_mdl_tbl <- list()
  }
  bivar_mdl_tbl[[names(cx_mdl_ft_a[i])]] <-
    gtsummary::tbl_regression(cx_mdl_ft_a[[i]],
                              exponentiate = T,
                              label = mdl_tab_lbl
                              ) %>%
    bold_labels() %>%
    add_significance_stars(hide_ci = F, hide_se = T)
}

# stack and export preped table
tbl_stack(bivar_mdl_tbl, group_header = NULL) %>% 
  as_gt() %>% 
  gt::tab_header(title = gt::md("Survival Models of CPR Registration: Unadjusted Hazard Ratios")) %>% 
  #convert output to a gt table for onward saving.
  gt::gtsave("outputs\\coxph_bivar_mdl_tbl_all.docx")


# print only fitted formula fit for documentation
for(i in seq_along(cx_mdl_ft_a)) {

  #   bivariate formula
  frml_a <- as.formula(paste0("Surv(c_age_1st_evnt, c_frst_cpr) ~ ", covar_all[[i]]))
  
  cat("\nModel", names(cx_mdl_ft_a[i]), "\n")
  print(frml_a)
}

```

## Multivariate demog Models
Child ethnicity is prefered to mothers ethnicity because a mother can have children with different ethnicities if for examples she's had more than one husbands.
```{r}
ehr_cpr_cens11_cohrt %>% glimpse

# Demographic Models
covar_demo2a <- c(
    "c_sex",
    "c_ethn",
    "m_age_cat",
    "m_edu",
    "m_hh_child_depndt",
    "m_hh_adult",
    "m_mar_status", #**
    # "m_uk_stay_len",
    "m_hh_deprv_no" #**
    )

# swaped "m_hh_deprv_no" with "m_hh_deprv_dms"
covar_demo2b <- c(
    "c_sex",
    "c_ethn",
    "m_age_cat",
    "m_edu",
    "m_hh_child_depndt",
    "m_hh_adult",
    "m_mar_status", #**
    # "m_uk_stay_len",
    "m_hh_deprv_dms" #**
    )

# named list with exposure of models with only demographic variables
cx_mdls_multi_vars <-
  list(demo2a=covar_demo2a,
       demo2b=covar_demo2b#,
       )
cx_mdls_multi_vars

# access name
names(cx_mdls_multi_vars[1])

# Fit models (based on prepped formula)
for(i in seq_along(cx_mdls_multi_vars)) {
  if (i == 1) {
    cx_mdl_multi <- list()
    mdl_params_multi <- list()
  }
  
  #   prep formula, fit models, process & output coefficients
  frml_multi <- as.formula(paste0(
    "Surv(c_age_1st_evnt, c_frst_cpr) ~ ",
    paste(cx_mdls_multi_vars[[i]], collapse = " + ")
  ))
  
  cat("\nModel", names(cx_mdls_multi_vars[i]), "\n")
  print(frml_multi)
  
  cx_mdl_multi[[names(cx_mdls_multi_vars[i])]] <- survival::coxph(formula = frml_multi,
                                       data    = ehr_cpr_cens11_cohrt,
                                       id      = c_alf_pe)
  
  cat("Processing model parameters...")
  # #     obtain and process counts per group
  vars_frq_cnt <-
    ehr_cpr_cens11_cohrt %>%
    select(c_age_1st_evnt, c_frst_cpr, cx_mdls_multi_vars[[i]]) %>% #added outcomes so that complete case analysis (via NA omit) mirrors actual model implementation
    na.omit() %>% #because analysis is often for complete cases
    gtsummary::tbl_summary(
      type = list(where(is.logical) ~ "categorical",
                  all_dichotomous() ~ "categorical"),
      missing_text = "Missing",
      statistic = list(
        all_continuous() ~ "{median} ({p25}, {p75})",
        all_categorical() ~ "{n} ({p}%)"
        # all_categorical(is.na) ~ "{p_miss}"
      ),
      digits = everything() ~ c(0, 2)
    ) %>%
    .$table_body %>%
    filter(row_type == "level") %>%
    mutate(Parameter = paste0(variable, label)) %>%
    select(Parameter, variable, label, n_perc = stat_0)
  
  #     obtain and process model parameters
  mdl_params_multi[[names(cx_mdls_multi_vars[i])]] <-
    model_parameters(
      cx_mdl_multi[[i]],
      #18, 9
      exponentiate = T,
      include_info = T,
      pretty_names = T,
      include_reference = T
    ) %>%
    left_join(vars_frq_cnt, by = "Parameter") %>%
    select(-c(Parameter, df_error)) %>% #drop no longer needed join column
    relocate(variable, label)
  
  if (i == length(cx_mdls_multi_vars)) {
    # lastly, save list of multiple models to xlsx workbook file in batch
    # save list of named dataframes as xlsx with one df per sheet
    writexl::write_xlsx(mdl_params_multi, path = "outputs\\multivar_coxph_cpr_demo.xlsx")
  
    # tabulate the performance of multiple models
  compare_performance(cx_mdl_multi) %>% 
  write_csv(file = "outputs\\multivar_coxph_cpr_demo_perf.csv")
    }
}

# model diagnosis
check_model(cx_mdl_multi[[1]], residual_type
  = "normal")

# tabulate models
for(i in seq_along(cx_mdl_multi)) {
  if (i == 1) {
    demog_mdl_tbl <- list()
  }
  demog_mdl_tbl[[names(cx_mdl_multi[i])]] <-
    gtsummary::tbl_regression(cx_mdl_multi[[i]],
                              exponentiate = T,
                              label = mdl_tab_lbl
                              ) %>%
    bold_labels() %>%
    add_significance_stars(hide_ci = F, hide_se = T)
}

# stack and export preped table
tbl_merge(demog_mdl_tbl, tab_spanner = paste("Model", seq.int(1,length(demog_mdl_tbl)))) %>% 
  as_gt() %>% 
  gt::tab_header(title = gt::md("Survival Models of CPR Registration: Demography adjusted Hazard Ratios")) %>% 
  #convert output to a gt table for onward saving.
  gt::gtsave("outputs\\coxph_demog_mdl_tbl_all.docx")

# print only fitted formula fit for documentation
for(i in seq_along(cx_mdls_multi_vars)) {
    #   prep formula, fit models, process & output coefficients
  frml_multi <- as.formula(paste0(
    "Surv(c_age_1st_evnt, c_frst_cpr) ~ ",
    paste(cx_mdls_multi_vars[[i]], collapse = " + ")
  ))
  
  cat("\nModel", names(cx_mdls_multi_vars[i]), "\n")
  print(frml_multi)
}
```
**demographic Models fit**

Model demo2a: 
Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + 
    m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_no

Model demo2b: 
Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + 
    m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_dms


## Multivariate full Models

I take Demographic models 2a & 2b and add all perinatal factors to each. Demog models

```{r}

ehr_cpr_cens11_cohrt %>% 
  glimpse

# full models
covar_full2a <- c(
    "c_sex",
    "c_ethn",
    "m_age_cat",
    "m_edu",
    "m_hh_child_depndt",
    "m_hh_adult",
    "m_mar_status", #**
    # "m_uk_stay_len",
    "m_hh_deprv_no", #**
        "c_cgm",
    "c_ptb",
    "c_bw",
    "m_alc",
    "m_lrn",
    "m_mnt",
    "m_sbm",
    "m_smk"
    )

# swaped "m_hh_deprv_no" with "m_hh_deprv_dms"
covar_full2b <- c(
    "c_sex",
    "c_ethn",
    "m_age_cat",
    "m_edu",
    "m_hh_child_depndt",
    "m_hh_adult",
    "m_mar_status", #**
    # "m_uk_stay_len",
    "m_hh_deprv_dms", #**
        "c_cgm",
    "c_ptb",
    "c_bw",
    "m_alc",
    "m_lrn",
    "m_mnt",
    "m_sbm",
    "m_smk"
    )

# named list with exposure of models with only demographic variables
cx_mdls_multi_full <-
  list(full_2a=covar_full2a,
       full_2b=covar_full2b#,
       # full_2c=covar_full2c,
       # full_2d=covar_full2d
       )

cx_mdls_multi_full

# access name
names(cx_mdls_multi_full[2])

# make formula
for(i in seq_along(cx_mdls_multi_full)) {
  if (i == 1) {
    cx_mdl_multi_full <- list()
    mdl_params_multi_full <- list()
    ehr_cpr_cens11_cohrt <- as.data.frame(ehr_cpr_cens11_cohrt)
  }
  
  #   prep formula, fit models, process & output coefficients
  frml_multi_full <- as.formula(paste0(
    "Surv(c_age_1st_evnt, c_frst_cpr) ~ ",
    paste(cx_mdls_multi_full[[i]], collapse = " + ")
  ))
  
  cat("\nModel", names(cx_mdls_multi_full[i]), "\n")
  print(frml_multi_full)
  
  cx_mdl_multi_full[[names(cx_mdls_multi_full[i])]] <-
    survival::coxph(formula = frml_multi_full,
                    data    = ehr_cpr_cens11_cohrt,
                    id      = c_alf_pe)
  
  cat("Processing model parameters...")
  # #     obtain and process counts per group
  vars_frq_cnt_full <-
    ehr_cpr_cens11_cohrt %>%
    select(c_age_1st_evnt, c_frst_cpr, cx_mdls_multi_full[[i]]) %>% #added outcomes so that complete case analysis (via NA omit) mirrors actual model implementation
    na.omit() %>% #because analysis is often for complete cases
    gtsummary::tbl_summary(
      type = list(
        where(is.logical) ~ "categorical",
        all_dichotomous() ~ "categorical"
      ),
      missing_text = "Missing",
      statistic = list(
        all_continuous() ~ "{median} ({p25}, {p75})",
        all_categorical() ~ "{n} ({p}%)"
        # all_categorical(is.na) ~ "{p_miss}"
      ),
      digits = everything() ~ c(0, 2)
    ) %>%
    .$table_body %>%
    filter(row_type == "level") %>%
    mutate(Parameter = paste0(variable, label) %>% str_squish()) %>%
    select(Parameter, variable, label, n_perc = stat_0) #%>%
  # View("cnt_full")
  
  #     obtain and process model parameters
  mdl_params_multi_full[[names(cx_mdls_multi_full[i])]] <-
    model_parameters(
      cx_mdl_multi_full[[i]],
      #18, 9
      exponentiate = T,
      include_info = T,
      pretty_names = T,
      include_reference = T
    ) %>%
    mutate(Parameter = str_squish(Parameter)) %>%
    left_join(vars_frq_cnt_full, by = "Parameter") %>%
    select(-c(z, df_error)) %>%
    relocate(variable, label)
  
  if (i == length(cx_mdls_multi_full)) {
    # lastly, save list of multiple models to xlsx workbook file in batch
    # save list of named dataframes as xlsx with one df per sheet
    writexl::write_xlsx(mdl_params_multi_full, path = "outputs\\multivar_coxph_cpr_full.xlsx")
    
    # tabulate performance of multiple models
  compare_performance(cx_mdl_multi_full) %>% 
  write_csv(file = "outputs\\multivar_coxph_full_cpr_perf.csv")
  # This shows that model 2 hasa the greatest explanatory power.
  }
}

# model diagnosis
check_model(cx_mdl_multi_full[[1]], residual_type
  = "normal")

# tabulate fully adjusted models
for(i in seq_along(cx_mdl_multi_full)) {
  if (i == 1) {
    fll_adj_mdl_tbl <- list()
  }
  fll_adj_mdl_tbl[[names(cx_mdl_multi_full[i])]] <-
    gtsummary::tbl_regression(cx_mdl_multi_full[[i]],
                              exponentiate = T,
                              label = mdl_tab_lbl
                              ) %>%
    bold_labels() %>%
    add_significance_stars(hide_ci = F, hide_se = T)
}

# stack and export preped table
tbl_merge(fll_adj_mdl_tbl, tab_spanner = paste("Model", seq.int(1,length(fll_adj_mdl_tbl)))) %>% 
  as_gt() %>% 
  gt::tab_header(title = gt::md("Survival Models of CPR Registration: Fully adjusted Hazard Ratios")) %>% 
  #convert output to a gt table for onward saving.
  gt::gtsave("outputs\\coxph_full_adj_mdl_tbl_all.docx")


# print only formula to fit for documentation
for(i in seq_along(cx_mdls_multi_full)) {
    #   prep formula, fit models, process & output coefficients
  frml_multi <- as.formula(paste0(
    "Surv(c_age_1st_evnt, c_frst_cpr) ~ ",
    paste(cx_mdls_multi_full[[i]], collapse = " + ")
  ))
  
  cat("\nModel", names(cx_mdls_multi_full[i]), "\n")
  print(frml_multi)
}

```
Model full_2a: 
Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + 
    m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_no + 
    c_cgm + c_ptb + c_bw + m_alc + m_lrn + m_mnt + m_sbm + m_smk

Model full_2b: 
Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + 
    m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_dms + 
    c_cgm + c_ptb + c_bw + m_alc + m_lrn + m_mnt + m_sbm + m_smk

## Model tabulations (focused on main exposures)
```{r}
# fitted model objects (named lists)
## Unadjusted models
names(cx_mdl_ft_a)

## demography adjusted models
names(cx_mdl_multi)

## fully adjusted models
names(cx_mdl_multi_full)

# tabulation 1
# m_hh_deprv_no + demo2a + full_2a
mdl_tbl_1 <- list()
mdl_tbl_1[["Unadjusted"]] <- 
gtsummary::tbl_regression(cx_mdl_ft_a[["m_hh_deprv_no"]],
                          exponentiate = T,
                          label = mdl_tab_lbl) %>% 
  bold_labels() %>% 
  add_significance_stars(hide_ci = F, hide_se = T) #%>%

# mdl_tbl_1[["Unadjusted"]]

mdl_tbl_1[["aHR 1"]] <- 
gtsummary::tbl_regression(cx_mdl_multi[["demo2a"]],
                          exponentiate = T,
                          label = mdl_tab_lbl) %>% 
  bold_labels() %>%
  add_significance_stars(hide_ci = F, hide_se = T) #%>% 
  # add_global_p(test = "Wald") %>% 

mdl_tbl_1[["aHR 2"]] <- 
gtsummary::tbl_regression(cx_mdl_multi_full[["full_2a"]],
                          exponentiate = T,
                          label = mdl_tab_lbl) %>% 
  bold_labels() %>%
  add_significance_stars(hide_ci = F, hide_se = T) %>% 
  # add_global_p(test = "Wald") %>% 
  add_n(location = "level") #%>% 
  # add_nevent(location = "level") #removed to motigate small numbers

mdl_tbl_1_all <- 
  tbl_merge(
    tbls = mdl_tbl_1,
    tab_spanner = c("**Unadjusted**",
                    "**Demography adjusted**",
                    "**Fully adjusted**"),
    quiet = T
  ) %>% 
  as_gt() %>% #convert output to a gt table for onward saving.
  gt::tab_header(title = gt::md("Progressively Adjusted Survival Models of CPR Registration (A): Focusing on HH Deprivation Count")) %>% 
  gt::gtsave("outputs\\coxph_mdl_tbl_1_all.docx")

# tabulation 3
# m_hh_deprv_dms + demo2b + full_2b
mdl_tbl_3 <- list()
mdl_tbl_3[["Unadjusted"]] <-
  gtsummary::tbl_regression(cx_mdl_ft_a[["m_hh_deprv_dms"]],
                            exponentiate = T,
                            label = mdl_tab_lbl) %>%
  bold_labels() %>%
  add_significance_stars(hide_ci = F, hide_se = T)

mdl_tbl_3[["aHR 1"]] <-
  gtsummary::tbl_regression(cx_mdl_multi[["demo2b"]],
                            exponentiate = T,
                            label = mdl_tab_lbl) %>%
  bold_labels() %>%
  add_significance_stars(hide_ci = F, hide_se = T)

mdl_tbl_3[["aHR 2"]] <-
  gtsummary::tbl_regression(cx_mdl_multi_full[["full_2b"]],
                            exponentiate = T,
                            label = mdl_tab_lbl) %>%
  bold_labels() %>%
  add_significance_stars(hide_ci = F, hide_se = T) %>%
  add_n(location = "level") #%>%
  # add_nevent(location = "level")

# merge tables and export as docx
mdl_tbl_3_all <-
  tbl_merge(
    tbls = mdl_tbl_3,
    tab_spanner = c(
      "**Unadjusted**",
      "**Demography adjusted**",
      "**Fully adjusted**"
    ),
    quiet = T
  ) %>%
  as_gt() %>% #convert output to a gt table for onward saving.
    gt::tab_header(title = gt::md("Progressively Adjusted Survival Models of CPR Registration (A): Focusing on HH Deprivation Dimensions")) %>%
  gt::gtsave("outputs\\coxph_mdl_tbl_3_all.docx")

# gc()
# save.image()

```

# Diagnostics of the Cox PH Models

## Bivariate models
```{r}

# cx_mdl_ft_a

length(cx_mdl_ft_a)

# model names
names(cx_mdl_ft_a)

# drop lenght of stay in uk from this batch for easy plotting
length(cx_mdl_ft_a[names(cx_mdl_ft_a) != "m_uk_stay_len"])

# covariate/exposure names
covar_all[[1]]

# test proportional hazard assumption
for(i in seq_along(cx_mdl_ft_a)) {
  if (i == 1) {
    # bivar_mdl_plt_lst <- list()
    bivar_zph_lvl_tbl_lst <- list()
    bivar_zph_plt_lst <- list()
    
  }
  
  cat("\nModel", names(cx_mdl_ft_a[i]), "diagnostics...\n")

  ftst <- cox.zph(cx_mdl_ft_a[[i]])
  # ftst
  
  ## for each category/level
  ftst_lvl <- cox.zph(cx_mdl_ft_a[[i]],
                    terms = F)
  # ftst_cat

  cat("\tProcessing parameters from model diagnosis...\n")
  # #     obtain and process counts per group
  vars_frq_cnt_full <-
    ehr_cpr_cens11_cohrt %>%
    select(c_age_1st_evnt, c_frst_cpr, covar_all[[i]]) %>% #added outcomes so that complete case analysis (via NA omit) mirrors actual model implementation
    na.omit() %>% #because analysis is often for complete cases
    gtsummary::tbl_summary(
      type = list(
        where(is.logical) ~ "categorical",
        all_dichotomous() ~ "categorical"
      ),
      missing_text = "Missing",
      statistic = list(
        all_continuous() ~ "{median} ({p25}, {p75})",
        all_categorical() ~ "{n} ({p}%)"
        # all_categorical(is.na) ~ "{p_miss}"
      ),
      digits = everything() ~ c(0, 2)
    ) %>%
    .$table_body %>%
    filter(row_type == "level") %>%
    mutate(Parameter = paste0(variable, label) %>% str_squish()) %>%
    select(Parameter, variable, label, n_perc = stat_0)

  ## per level
  bivar_zph_lvl_tbl_lst[[names(cx_mdl_ft_a[i])]] <-
    vars_frq_cnt_full %>%
    full_join(ftst_lvl$table %>%
                as_tibble(rownames = "Parameter"), by = "Parameter") %>%
    relocate(variable, label)
  
  cat("\tProducing model diagnosis...\n")
  
  # this produces a list of plots (based on the number of covariates)
  mdl_ftst <- ggcoxzph(ftst, resid = F)
  
#   obtain plot with inner list
  bivar_zph_plt_lst[[names(cx_mdl_ft_a[i])]] <- 
    mdl_ftst[[1]]
  
  if (i == length(cx_mdl_ft_a)) {

    cat("Saving parameters of all model diagnosis...\n")

    writexl::write_xlsx(bivar_zph_lvl_tbl_lst, path = "outputs\\bivar_cpr_zph_lvl_tbls.xlsx")
    
    cat("Consolidating & saving all plots of model diagnosis...\n")
    #   consolidate/mosaic all plots of bivar coxph models

  bivar_zph_plt <-
    wrap_plots(bivar_zph_plt_lst[names(bivar_zph_plt_lst) != "m_uk_stay_len"],
               ncol = 4,
               nrow = 5,
               tag_level = 'keep',
    ) +
    plot_annotation(
      subtitle = paste0("Bivariate model ", "diagnostics"),
      tag_levels = 'A'
    )

  ggsave(
    filename = paste0("outputs\\zph_bivar_","plts.png"),
    plot = bivar_zph_plt,
    height = 50,
    width = 50,
    units = 'cm',
    dpi = 600
  )
  
#     cleanup workspace
    rm(ftst, ftst_lvl, vars_frq_cnt_full, mdl_ftst)
  }
}

gc()
save.image()
```

## Multivariate Demog models

In these results, variables whose p values are significant (i.e. = 0.1) violate the
HR assumption of being constant through time. They are therefore better fit using
discrete time or time varying models since their HR varies with time.

```{r}

names(cx_mdl_multi)
cx_mdls_multi_vars[[1]]

# test proportional hazard assumption
for(i in seq_along(cx_mdl_multi)) {
  if (i == 1) {
    demo_zph_tbl_lst <- list()
    demo_zph_lvl_tbl_lst <- list()
    demo_zph_plt_lst <- list()
    
  }
  
  cat("\nModel", names(cx_mdl_multi[i]), "diagnostics...\n")
  #   perform test (on each model)
   ## for each variable
  ftst <- cox.zph(cx_mdl_multi[[i]])
  # ftst
  
  ## for each category/level
  ftst_lvl <- cox.zph(cx_mdl_multi[[i]],
                    terms = F)
  # ftst_cat

  cat("\tProcessing parameters from model diagnosis...\n")
  # #     obtain and process counts per group
  vars_frq_cnt_full <-
    ehr_cpr_cens11_cohrt %>%
    select(c_age_1st_evnt, c_frst_cpr, cx_mdls_multi_vars[[i]]) %>% #added outcomes so that complete case analysis (via NA omit) mirrors actual model implementation
    na.omit() %>% #because analysis is often for complete cases
    gtsummary::tbl_summary(
      type = list(
        where(is.logical) ~ "categorical",
        all_dichotomous() ~ "categorical"
      ),
      missing_text = "Missing",
      statistic = list(
        all_continuous() ~ "{median} ({p25}, {p75})",
        all_categorical() ~ "{n} ({p}%)"
        # all_categorical(is.na) ~ "{p_miss}"
      ),
      digits = everything() ~ c(0, 2)
    ) %>%
    .$table_body %>%
    filter(row_type == "level") %>%
    mutate(Parameter = paste0(variable, label) %>% str_squish()) %>%
    select(Parameter, variable, label, n_perc = stat_0)

  #   extract tables from test object
  ## per variable
  demo_zph_tbl_lst[[names(cx_mdl_multi[i])]] <-
    ftst$table %>%
    as_tibble(rownames = "Variable")
  
  ## per level
  demo_zph_lvl_tbl_lst[[names(cx_mdl_multi[i])]] <-
    ftst_lvl$table %>%
    as_tibble(rownames = "Parameter") %>% 
      left_join(vars_frq_cnt_full, by = "Parameter") %>%
    relocate(variable, label)
  
  cat("\tProducing, consolidating & saving plots of model diagnosis...\n")
  # this produces a list of plots (based on the number of covariates)
  mdl_ftst <- ggcoxzph(ftst, resid = F)

  #   consolidate/mosaic all plots of each model
  demo_zph_plt_lst[[names(cx_mdl_multi[i])]] <-
    wrap_plots(mdl_ftst,
               ncol = 2,
               nrow = 4,
               tag_level = 'keep',
    ) +
    plot_annotation(
      subtitle = paste0("Demographic model ", as.character(i), " diagnostics"),
      tag_levels = 'A'
    )

  ggsave(
    filename = paste0("outputs\\zph_demo_", names(cx_mdl_multi[i]), "_plts.png"),
    plot = demo_zph_plt_lst[[i]],
    height = 40,
    width = 25,
    units = 'cm',
    dpi = 600
  )
  
  if (i == length(cx_mdl_multi)) {
    cat("Saving parameters of all model diagnosis...\n")
    # lastly, save list of multiple models to xlsx workbook file in batch
    # save list of named dataframes as xlsx with one df per sheet
    writexl::write_xlsx(demo_zph_tbl_lst, path = "outputs\\multivar_demo_cpr_zph_tbls.xlsx")
    writexl::write_xlsx(demo_zph_lvl_tbl_lst, path = "outputs\\multivar_demo_cpr_zph_lvl_tbls.xlsx")
    
#     cleanup workspace
    rm(ftst, ftst_lvl, vars_frq_cnt_full, mdl_ftst)
  }
}


```

## Multivariate full models

In these results, variables whose p values a significant (i.e. \<= 0.1) violate the
HR assumption of being constant through time. They are therefore better fit using
discrete time or time varying models since their HR varies with time.

```{r}
# install.packages("ggplotify", dependencies = T)
# library(ggplotify)

names(cx_mdl_multi_full[1])

# test proportional hazard assumption
for(i in seq_along(cx_mdl_multi_full)) {
  if (i == 1) {
    full_zph_tbl_lst <- list()
    full_zph_lvl_tbl_lst <- list()
    full_zph_plt_lst <- list()
    
  }
  
  cat("\nModel", names(cx_mdls_multi_full[i]), "diagnostics...\n")
  #   perform test (on each model)
   ## for each variable
  ftst <- cox.zph(cx_mdl_multi_full[[i]])
  # ftst
  
  ## for each category/level
  ftst_lvl <- cox.zph(cx_mdl_multi_full[[i]],
                    terms = F)
  # ftst_cat

  cat("\tProcessing parameters from model diagnosis...\n")
  # #     obtain and process counts per group
  vars_frq_cnt_full <-
    ehr_cpr_cens11_cohrt %>%
    select(c_age_1st_evnt, c_frst_cpr, cx_mdls_multi_full[[i]]) %>% #added outcomes so that complete case analysis (via NA omit) mirrors actual model implementation
    na.omit() %>% #because analysis is often for complete cases
    gtsummary::tbl_summary(
      type = list(
        where(is.logical) ~ "categorical",
        all_dichotomous() ~ "categorical"
      ),
      missing_text = "Missing",
      statistic = list(
        all_continuous() ~ "{median} ({p25}, {p75})",
        all_categorical() ~ "{n} ({p}%)"
        # all_categorical(is.na) ~ "{p_miss}"
      ),
      digits = everything() ~ c(0, 2)
    ) %>%
    .$table_body %>%
    filter(row_type == "level") %>%
    mutate(Parameter = paste0(variable, label) %>% str_squish()) %>%
    select(Parameter, variable, label, n_perc = stat_0)

  #   extract tables from test object
  ## per variable
  full_zph_tbl_lst[[names(cx_mdl_multi_full[i])]] <-
    ftst$table %>%
    as_tibble(rownames = "Variable")
  
  ## per level
  full_zph_lvl_tbl_lst[[names(cx_mdl_multi_full[i])]] <-
    ftst_lvl$table %>%
    as_tibble(rownames = "Parameter") %>% 
      left_join(vars_frq_cnt_full, by = "Parameter") %>%
    relocate(variable, label)
  
  cat("\tProducing, consolidating & saving plots of model diagnosis...\n")
  # this produces a list of plots (based on the number of covariates)
  mdl_ftst <- ggcoxzph(ftst, resid = F)

  #   consolidate/mosaic all plots of each model
  full_zph_plt_lst[[names(cx_mdl_multi_full[i])]] <-
    wrap_plots(mdl_ftst,
               ncol = 4,
               nrow = 4,
               tag_level = 'keep',
    ) +
    plot_annotation(
      subtitle = paste0("Full model ", as.character(i), " diagnostics"),
      tag_levels = 'A'
    )

  ggsave(
    filename = paste0("outputs\\zph_", names(cx_mdl_multi_full[i]), "_plts.png"),
    plot = full_zph_plt_lst[[i]],
    height = 40,
    width = 50,
    units = 'cm',
    dpi = 600
  )
  
  if (i == length(cx_mdl_multi_full)) {
    cat("\tSaving parameters of model diagnosis...\n")
    # lastly, save list of multiple models to xlsx workbook file in batch
    # save list of named dataframes as xlsx with one df per sheet
    writexl::write_xlsx(full_zph_tbl_lst, path = "outputs\\multivar_full_cpr_zph_tbls.xlsx")
    writexl::write_xlsx(full_zph_lvl_tbl_lst, path = "outputs\\multivar_full_cpr_zph_lvl_tbls.xlsx")
    
#     cleanup workspace
    rm(ftst, ftst_lvl, vars_frq_cnt_full, mdl_ftst)
  }
}

```

# Attributable Fractions
## Unadjusted (bivariate) Models
```{r}
# var list
length(covar_all)

# fitted coxph model
cx_mdl_ft_a

# inner loop for each model
for (j in seq_along(covar_all)) {
  if (j == 1) {
    var_no <- length(covar_all)
    cat("No. of variables:", var_no, "\n")
    
    # followup years for AF computation
    follow_up_yr <- seq.int(1, 11)
    # follow_up_yr
    cx_mdl_unadj_af <- list()
    # mdl_params_multi <- list()
    
    #     process AF table for neatness
    tidy_af <- function(af_objt) {
      af_objt2 <- summary(af_objt)
      # af_objt
      
      af_tbl <-
        dplyr::bind_cols(
          age_yr = af_objt2$times,
          af_objt2$AF,
          af_objt2$confidence.interval,
          .name_repair = janitor::make_clean_names
        )
      
      return(af_tbl)
    }
    
    # create indicator variables of levels relevant to the AF computation &
    #    coerce tibble to raw dataframe for the AFcoxph function
    ehr_cpr_cens11_cohrt_indctr <-
      ehr_cpr_cens11_cohrt %>%
      fastDummies::dummy_cols(
        select_columns = covar_all,
        remove_first_dummy = F,
        ignore_na = T
      ) %>%
      mutate(c_cpr_evnt = as.integer(c_cpr_evnt)) %>%
      as.data.frame() #%>%
      # glimpse
    
  }
  
  cat("\nAttributable Fractions for '",
      names(covar_all[j]),
      "'\n")
  
  #   extract variable
  af_var <- covar_all[j]
  
  fct_lvls_x <-
    ehr_cpr_cens11_cohrt_indctr %>%
    #   drop original variable names
    select(-any_of(names(ehr_cpr_cens11_cohrt))) %>% #glimpse
    #   levels for var j
    select(starts_with(af_var)) %>%
    names
  
  cat("\n  Variable:", af_var, "\n")

  for (k in seq_along(fct_lvls_x)) {
    cat("\tLevel:", fct_lvls_x[k], "\n")
    # replace original variable in formular j with indicator variable k in turn
    af_fmla <-
      cx_mdl_ft_a[[j]]$formula %>%
      deparse1() %>%
      str_replace(af_var, paste0("`", fct_lvls_x[k], "`")) %>%
      #   replace logical with binary outcome variable
      str_replace("c_frst_cpr", "c_cpr_evnt") %>%
      as.formula()
    
    # print(af_fmla)
    cat("\t", deparse1(af_fmla), "\n")
    
    # fit coxph for AF
    cx_fit_1 <- coxph(formula = af_fmla,
                      data = ehr_cpr_cens11_cohrt_indctr,
                      ties = "breslow")
    
    #       helps skip AF calculation which failt
    try({
      # attributeable fraction calculation
      af_cx_fit_1 <-
        AF::AFcoxph(
          object = cx_fit_1,
          data = ehr_cpr_cens11_cohrt_indctr,
          exposure = paste0(fct_lvls_x[k]),
          # exposure = paste0(fct_lvls_x[1]),
          times = follow_up_yr,
          #if not, numerous calculations are made at 1 decimal place time points
          clusterid = "c_alf_pe"
        )
      
      #   concat to neat table
      cx_mdl_unadj_af[[paste0(covar_all[j], ".", fct_lvls_x[k])]] <-
        tidy_af(af_cx_fit_1)
      # af_cx_fit_1_tbl
      
    })

  }
      
  if (j == var_no) {
    #     stack AF stats of all vars in this model
    cx_mdl_unadj_af_all <-
      bind_rows(cx_mdl_unadj_af, .id = "mdl_var")
    
          # save list of named dataframes as xlsx with one df per sheet
      writexl::write_xlsx(cx_mdl_unadj_af_all, path = "outputs\\cx_mdl_unadj_af_tbls_all.xlsx")
  }
}

save.image()
```

## Demographic Models

AF computed for demographic models a & b only

```{r}

# make formula
for(i in seq_along(cx_mdls_multi_vars)) {
  if (i == 1) {
    cx_mdl_demo_af_all <- list()
    # mdl_params_multi <- list()
    
    #     process AF table for neatness
    tidy_af <- function(af_objt) {
      af_objt2 <- summary(af_objt)
      # af_objt
      
      af_tbl <-
        dplyr::bind_cols(
          age_yr = af_objt2$times,
          af_objt2$AF,
          af_objt2$confidence.interval,
          .name_repair = janitor::make_clean_names
        )
      
      return(af_tbl)
    }
    
    # flatten all list to vector, and make unique
    demo_vars_all <- 
      cx_mdls_multi_vars %>% 
      list_c() %>% 
      unique()
    
    # create indicator variables of levels relevant to the AF computation &
    #    coerce tibble to raw dataframe for the AFcoxph function
    ehr_cpr_cens11_cohrt_indctr <-
      ehr_cpr_cens11_cohrt %>%
      fastDummies::dummy_cols(
        select_columns = demo_vars_all,
        remove_first_dummy = F,
        ignore_na = T
      ) %>%
      mutate(c_cpr_evnt = as.integer(c_cpr_evnt)) %>%
      as.data.frame() #%>%
      # glimpse
    
    #     track analysis progress
    mdl_cnt <- length(cx_mdls_multi_vars)
    cat("No. of Models:", mdl_cnt, "\n")
  }
  
  cat("\n----------------------------------------\n")
  mdl_nme <- names(cx_mdls_multi_vars[i])
  cat("Model Name:", mdl_nme, "\n")
  # }
  
  # inner loop for each model
  for (j in seq_along(cx_mdls_multi_vars[[i]])) {
    if (j == 1) {
      var_no <- length(cx_mdls_multi_vars[[i]])
      cat("No. of variables:", var_no, "\n")
      
      # followup years for AF computation
      follow_up_yr <- seq.int(1, 11)
      # follow_up_yr
      cx_mdl_multi_demo_af <- list()
      # mdl_params_multi <- list()
          
      cat("\nAttributable Fractions for model",
          names(cx_mdls_multi_vars[i]),
          "\n")
    }
    
    #   extract variable
    af_var <- cx_mdls_multi_vars[[i]][j]
    
    fct_lvls_x <-
      ehr_cpr_cens11_cohrt_indctr %>%
      #   drop original variable names
      select(-any_of(names(ehr_cpr_cens11_cohrt))) %>% #glimpse
      #   levels for var i
      select(starts_with(af_var)) %>%
      names
    
    cat("\n  Variable:", af_var, "\n")
    
    for (k in seq_along(fct_lvls_x)) {
      cat("\tLevel:", fct_lvls_x[k], "\n")
      # replace original variable in formular i with indicator variable k in turn
      af_fmla <-
        cx_mdl_multi[[i]]$formula %>%
        deparse1() %>%
        str_replace(af_var, paste0("`", fct_lvls_x[k], "`")) %>%
        #   replace logical with binary outcome variable
        str_replace("c_frst_cpr", "c_cpr_evnt") %>%
        as.formula()
      
      # print(af_fmla)
      cat("\t", deparse1(af_fmla), "\n")
      
      # fit coxph for AF
      cx_fit_1 <- coxph(formula = af_fmla,
                        data = ehr_cpr_cens11_cohrt_indctr,
                        ties = "breslow")
      
      #       helps skip AF calculation which failt
      try({
        # attributeable fraction calculation
        af_cx_fit_1 <-
          AF::AFcoxph(
            object = cx_fit_1,
            data = ehr_cpr_cens11_cohrt_indctr,
            exposure = paste0(fct_lvls_x[k]),
            # exposure = paste0(fct_lvls_x[1]),
            times = follow_up_yr,
            #if not, numerous calculations are made at 1 decimal place time points
            clusterid = "c_alf_pe"
          )
        
        #   concat to neat table
        cx_mdl_multi_demo_af[[paste0(names(cx_mdls_multi_vars[i]), ".", fct_lvls_x[k])]] <-
          tidy_af(af_cx_fit_1)
        # af_cx_fit_1_tbl
        
      })
      
      if (j == var_no) {
        #     stack AF stats of all vars in this model
        cx_mdl_demo_af_all[[paste0(mdl_nme)]] <-
          bind_rows(cx_mdl_multi_demo_af, .id = "mdl_var")
      }
      
    }
    
  }
  
  if (i == mdl_cnt) {
      # save list of named dataframes as xlsx with one df per sheet
      writexl::write_xlsx(cx_mdl_demo_af_all, path = "outputs\\cx_mdl_demo_af_tbls_all.xlsx")
    }
}

gc()
save.image()

```

In some models, AF could not be calculated for some factors due to the following
types of errors, probably because those factors has too few cases:

Variable: c_sex Level: c_sex_M Surv(c_age_1st_evnt, c_cpr_evnt) ~ c_sex_M + c_ethn +
m_age_cat + m_edu + m_hh_child_depndt + m_hh_adult + m_fam_status + m_hh_deprv_no
Error in solve.default(vcov(object = object)) : Lapack routine dgesv: system is
exactly singular: U[26,26] = 0 In addition: Warning message: In coxph.fit(X, Y,
istrat, offset, init, control, weights = weights, : Loglik converged before variable
24 ; coefficient may be infinite. Level: c_sex_F Surv(c_age_1st_evnt, c_cpr_evnt) ~
c_sex_F + c_ethn + m_age_cat + m_edu + m_hh_child_depndt + m_hh_adult +
m_fam_status + m_hh_deprv_no Error in solve.default(vcov(object = object)) : Lapack
routine dgesv: system is exactly singular: U[26,26] = 0 In addition: Warning message:
In coxph.fit(X, Y, istrat, offset, init, control, weights = weights, : Loglik
converged before variable 24 ; coefficient may be infinite. Attributable Fractions
for model demo2c

An alternative to the AF approach for coxph, is to plot adjusted survival curves
using the ggcoxadjustedcurves function of survminer

## Full models

```{r}
# ehr_cpr_cens11_cohrt %>% glimpse

# flatten all list to vector, and make unique

cx_mdls_multi_full
# demo_vars_all <-
# cx_mdls_multi_vars %>%
# list_c() %>%
# unique()

# make formula

for (i in seq_along(cx_mdls_multi_full)) {
  if (i == 1) {
    cx_mdl_full_af_all <- list()
    # mdl_params_multi <- list()
    
    #     process AF table for neatness
    tidy_af <- function(af_objt) {
      af_objt2 <- summary(af_objt)
      # af_objt2
      
      af_tbl <-
        dplyr::bind_cols(
          age_yr = af_objt2$times,
          af_objt2$AF,
          af_objt2$confidence.interval,
          .name_repair = janitor::make_clean_names
        )
      
      return(af_tbl)
    }
    
#     flatten variable list of lists, and return unique variables
    full_vars_all <-
      cx_mdls_multi_full %>%
      list_c() %>%
      unique()
    
    # create indicator variables of levels relevant to the AF computation
    #     then coearce tibble a raw dataframe for the AFcoxph function
    ehr_cpr_cens11_cohrt_indctr_full <-
      ehr_cpr_cens11_cohrt %>%
      droplevels() %>%
      fastDummies::dummy_cols(
        select_columns = full_vars_all,
        remove_first_dummy = F,
        ignore_na = T
      ) %>%
      mutate(c_cpr_evnt = as.integer(c_cpr_evnt)) %>%
      as.data.frame()
    
    #     track analysis progress
    mdl_cnt <- length(cx_mdls_multi_full)
    cat("No. of Models:", mdl_cnt, "\n")
    
  }
  
  cat("\n----------------------------------------\n")
  
  mdl_nme <-
    names(cx_mdls_multi_full[i])
  cat("Model Name:", mdl_nme, "\n")
  
# }
  
  # inner loop for each model
  for (j in seq_along(cx_mdls_multi_full[[i]])) {
    if (j == 1) {
      
      var_no <- length(cx_mdls_multi_full[[i]])
      cat("No. of variables:", var_no, "\n")
      
      # followup years for AF computation
      follow_up_yr <- seq.int(1, 11)
      # follow_up_yr
      cx_mdl_multi_full_af <- list()
      # mdl_params_multi <- list()
      
    cat("\nAttributable Fractions for model",
        names(cx_mdls_multi_full[i]), "\n")
    }
    
    #   extract variable
    af_var <- cx_mdls_multi_full[[i]][j]
    
    fct_lvls_x <-
      ehr_cpr_cens11_cohrt_indctr_full %>%
      #   drop original variable names
      select(-any_of(names(ehr_cpr_cens11_cohrt))) %>% #glimpse
      #   levels for var i
      select(starts_with(af_var)) %>%
      names
    
    cat("\n  Variable:", af_var, "\n")
    
    for (k in seq_along(fct_lvls_x)) {
      cat("\tLevel:", fct_lvls_x[k], "\n")
      # replace original variable in formular i with indicator variable k in turn
      af_fmla <-
        cx_mdl_multi_full[[i]]$formula %>%
        deparse1() %>%
        str_replace(af_var, paste0("`", fct_lvls_x[k], "`")) %>%
        #   replace logical with binary outcome variable, and updated naming of outcomes
        # str_replace("frst_cpr", "c_cpr_evnt") %>%
        # str_replace("age_1st_evnt", "c_age_1st_evnt") %>%
        as.formula()
      
      # print(af_fmla)
      cat("\t", deparse1(af_fmla), "\n")
      
      # fit coxph for AF
      cx_fit_1 <- coxph(formula = af_fmla,
                        data = ehr_cpr_cens11_cohrt_indctr_full,
                        ties = "breslow")
      
      #       helps skip AF calculation which failt
      try({
        # attributeable fraction calculation
        af_cx_fit_1 <-
          AF::AFcoxph(
            object = cx_fit_1,
            data = ehr_cpr_cens11_cohrt_indctr_full,
            exposure = paste0(fct_lvls_x[k]),
            # exposure = paste0(fct_lvls_x[1]),
            times = follow_up_yr,
            #if not, numerous calculations are made at 1 decimal place time points
            clusterid = "c_alf_pe"
          )
        
        #   concat to neat table
        cx_mdl_multi_full_af[[paste0(names(cx_mdls_multi_full[i]), ".", fct_lvls_x[k])]] <-
          tidy_af(af_cx_fit_1)
        # af_cx_fit_1_tbl
        
      })
      
      if (j == var_no) {
        #     stack AF stats of all vars in this model
        cx_mdl_full_af_all[[paste0(mdl_nme)]] <-
          bind_rows(cx_mdl_multi_full_af, .id = "mdl_var")
      }
      
    }
    
  }
  
  if (i == mdl_cnt) {
    # save list of named dataframes as xlsx with one df per sheet
    writexl::write_xlsx(cx_mdl_full_af_all, path =
                          "outputs\\cx_mdl_full_af_tbls_all.xlsx")
  }
}

full_vars_all[1]

cx_mdl_full_af_all

gc()
save.image()

```
Only AF for full models a and b were calculated successfully for all variables. Models C and d calculated for only marital status. All other variables returned errors similar to the following and failed to calculate:
Level: c_sex_F 
	 Surv(c_age_1st_evnt, c_cpr_evnt) ~ c_sex_F + c_ethn + m_age_cat + m_edu + m_hh_child_depndt + m_hh_adult + m_fam_status + m_hh_deprv_no + c_cgm + c_ptb + c_bw + m_alc + m_lrn + m_mnt + m_sbm + m_smk 
Error in solve.default(vcov(object = object)) : 
  Lapack routine dgesv: system is exactly singular: U[26,26] = 0
In addition: Warning message:
In coxph.fit(X, Y, istrat, offset, init, control, weights = weights,  :
  Loglik converged before variable  24 ; coefficient may be infinite. 
  
It appears much of this problem might be from variables with variabes encoded unknown lining up to create collinearity (and then singularity problems)

# Attributable Fractions (graphPAF)

## PAF helper functions

Targeted graphPAF intervention helper for categorical/factor exposures

Purpose:
  Estimate the population attributable fraction (PAF) for hypothetical
  interventions that move one observed level of a factor exposure to a chosen
  reference/intervention level, while all other exposure levels remain as
  observed.

Example:
  For m_hh_deprv_no with levels 0, 1, 2, 3 and 4, setting ref_lvl = "0"
  estimates separate scenarios for 1 -> 0, 2 -> 0, 3 -> 0 and 4 -> 0.
```{r}

# Convert graphPAF::impact_fraction(..., ci = TRUE) output to a tidy dataframe.
tidy_graphpaf_if <- function(if_obj) {
  if ("time_vec" %in% names(if_obj)) {
    out <- tibble(
      Time = as.numeric(if_obj$time_vec),
      Estimated_PAF = as.numeric(if_obj$estimate),
      CI = as.character(if_obj$confidence_interval)
    )
  } else {
    out <- tibble(
      Time = NA_real_,
      Estimated_PAF = as.numeric(if_obj$estimate),
      CI = as.character(if_obj$confidence_interval)
    )
  }

  out %>%
    mutate(
      conf.low = as.numeric(str_match(CI, "\\(?\\s*([^,]+)\\s*,")[, 2]),
      conf.high = as.numeric(str_match(CI, ",\\s*([^\\)]+)\\s*\\)?")[, 2])
    )
}

# Estimate intervention PAFs for every eligible level of a factor exposure.
calc_deprv_level_paf <- function(ftd_mdl,
                                 data,
                                 fct_var,
                                 ref_lvl,
                                 excld_lvl = NULL,
                                 boot_rep = 100,
                                 t_vector = NULL,
                                 calculation_method = "D",
                                 ci_level = 0.95,
                                 ci_type = "norm",
                                 verbose = FALSE) {
  fct_var_chr <- as_name(ensym(fct_var))
  ref_lvl <- as.character(ref_lvl)
  excld_lvl <- as.character(excld_lvl)

  if (!is.data.frame(data)) {
    stop("`data` must be a dataframe containing the variables used to fit `ftd_mdl`.")
  }

  if (!fct_var_chr %in% names(data)) {
    stop("`fct_var` was not found in `data`.")
  }

  # graphPAF requires original and counterfactual datasets with matching
  # dimensions, column names and column classes. Work on a data.frame copy and
  # preserve row names used by fitted model objects.
  data <- as.data.frame(data)
  data_row_names <- row.names(data)

  data <- data %>%
    mutate("{fct_var_chr}" := fct_drop(as_factor(.data[[fct_var_chr]])))

  row.names(data) <- data_row_names

  fct_lvls <- levels(data[[fct_var_chr]])

  if (!ref_lvl %in% fct_lvls) {
    stop("`ref_lvl` must be one of the observed levels of `fct_var`.")
  }

  missing_exclusions <- setdiff(excld_lvl, fct_lvls)
  if (length(missing_exclusions) > 0) {
    stop("All values in `excld_lvl` must be observed levels of `fct_var`.")
  }

  levels_to_move <- setdiff(fct_lvls, c(ref_lvl, excld_lvl))

  if (length(levels_to_move) == 0) {
    stop("No eligible factor levels remain after excluding `ref_lvl` and `excld_lvl`.")
  }

  if (inherits(ftd_mdl, "coxph") && is.null(t_vector)) {
    stop("`t_vector` must be supplied when `ftd_mdl` is a Cox proportional hazards model.")
  }

  if (inherits(ftd_mdl, "coxph") && calculation_method != "D") {
    stop("For Cox proportional hazards models, graphPAF requires `calculation_method = \"D\"`.")
  }

  # graphPAF's bootstrap code expects the boot.ncpus option to be set.
  if (is.null(getOption("boot.ncpus"))) {
    options(boot.ncpus = 1)
  }

  map_dfr(levels_to_move, function(level_from) {
    # forcats::fct_recode() uses "new level" = "old level" syntax.
    # Here, one exposure level is moved to the reference/intervention level.
    recode_spec <- set_names(as.list(level_from), ref_lvl)

    new_data <- data %>%
      mutate(
        "{fct_var_chr}" := .data[[fct_var_chr]] %>%
          fct_recode(!!!recode_spec) %>%
          fct_expand(!!!fct_lvls) %>%
          fct_relevel(fct_lvls)
      )

    row.names(new_data) <- data_row_names

    if_obj <- graphPAF::impact_fraction(
      model = ftd_mdl,
      data = data,
      new_data = new_data,
      calculation_method = calculation_method,
      ci = TRUE,
      boot_rep = boot_rep,
      t_vector = t_vector,
      ci_level = ci_level,
      ci_type = ci_type,
      verbose = verbose
    )

    tidy_graphpaf_if(if_obj) %>%
      mutate(
        variable = fct_var_chr,
        level_from = as.character(level_from),
        level_to = ref_lvl,
        excluded_levels = if_else(
          length(excld_lvl) == 0,
          NA_character_,
          str_c(excld_lvl, collapse = "; ")
        ),
        scenario = str_c(fct_var_chr, ": ", level_from, " -> ", ref_lvl),
        .before = 1
      )
  })
}


# -------------------------------------------------------------------------
# Example 1: four separate interventions for household deprivation count.
# -------------------------------------------------------------------------

# deprv_no_unadj_cxph <-
#   survival::coxph(
#     formula = survival::Surv(c_age_1st_event, c_frst_cpr) ~ m_hh_deprv_no,
#     data = ehr_cpr_cens11_cohrt,
#     id = c_alf_pe
#   )
#
# follow_up_yr <- seq.int(0, 11)
#
# paf_deprv_no_to_zero <-
#   calc_deprv_level_paf(
#     ftd_mdl = deprv_no_unadj_cxph,
#     data = ehr_cpr_cens11_cohrt,
#     fct_var = m_hh_deprv_no,
#     ref_lvl = "0",
#     boot_rep = 500,
#     t_vector = follow_up_yr
#   )

# -------------------------------------------------------------------------
# Example 2: move only deprivation counts above 1 to level 1.
# -------------------------------------------------------------------------

# This estimates separate scenarios for 2 -> 1, 3 -> 1 and 4 -> 1. Level 0 is
# excluded so deprivation-free households are not counterfactually moved to 1.
#
# paf_deprv_gt1_to_one <-
#   calc_deprv_level_paf(
#     ftd_mdl = deprv_no_unadj_cxph,
#     data = ehr_cpr_cens11_cohrt,
#     fct_var = m_hh_deprv_no,
#     ref_lvl = "1",
#     excld_lvl = "0",
#     boot_rep = 500,
#     t_vector = follow_up_yr
#   )

# -------------------------------------------------------------------------
# Example 3: same function with a non-numeric factor exposure.
# -------------------------------------------------------------------------

# paf_deprv_dims_to_none <-
#   calc_deprv_level_paf(
#     ftd_mdl = deprv_dims_unadj_cxph,
#     data = ehr_cpr_cens11_cohrt,
#     fct_var = m_hh_deprv_dms,
#     ref_lvl = "None",
#     boot_rep = 500,
#     t_vector = follow_up_yr
#   )

# source further Counterfactual graphPAF helpers for household deprivation exposures
source("calc_deprv_counterfactual_paf_graphpaf.R")
# -------------------------------------------------------------------------
# Example workflow
# -------------------------------------------------------------------------
#
# paf_no_unadj <- calc_dprv_no_rdctn_paf(
#   ftd_mdl = deprv_no_unadj_cxph,
#   data = ehr_cpr_cens11_cohrt,
#   fct_var = m_hh_deprv_no,
#   dprv_rdctn_no = 1,
#   t_vector = seq.int(0, 11)
# )
#
# paf_no_plot <- plot_dprv_no_rdctn_paf(
#   paf_results = list(
#     "Unadjusted" = paf_no_unadj,
#     "Demography adjusted" = paf_no_demography,
#     "Fully adjusted" = paf_no_full
#   ),
#   time_point = 6
# )
#
# paf_dms_unadj <- calc_dprv_dms_rmvl_paf(
#   ftd_mdl = deprv_dms_unadj_cxph,
#   data = ehr_cpr_cens11_cohrt,
#   fct_var = m_hh_deprv_dms,
#   t_vector = seq.int(0, 11)
# )
#
# paf_dms_plot <- plot_dprv_dms_rmvl_paf(
#   paf_results = list("Unadjusted" = paf_dms_unadj),
#   time_point = 6
# )

```

## Deprivation combinations
### Unadjusted (bivariate) Models
```{r}

# use counts of deprivation

# use combinations of deprivation

deprv_dms_unadj_cxph <- survival::coxph(formula = Surv(c_age_1st_evnt, c_frst_cpr) ~ m_hh_deprv_dms,
                                    data    = ehr_cpr_cens11_cohrt,
                                    id      = c_alf_pe)
summary(deprv_dms_unadj_cxph)


ehr_cpr_cens11_cohrt |> 
  frq(m_hh_deprv_dms)

# compute population AF: set each dimension to none
deprv_dms_to_none <-
  paf_deprv_no_to_zero <-
    calc_deprv_level_paf(
      ftd_mdl = deprv_dms_unadj_cxph,
      data = ehr_cpr_cens11_cohrt,
      fct_var = m_hh_deprv_dms,
      output_file = "paf_dms_2_none_unadj.csv",
      ref_lvl = "None",
      boot_rep = 100,
      t_vector = follow_up_yr
    )
deprv_dms_to_none |> glimpse()

# compute population AF: remove each unitary dimension in turn
tic()
Sys.time()
paf_dms_unadj <- calc_dprv_dms_rmvl_paf(
  ftd_mdl = deprv_dms_unadj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_dms,
  t_vector = seq.int(0, 11),
  output_file = "paf_dms_unadj.csv"
)
Sys.time()
toc() #1162.31 sec elapsed (20 mins)
paf_dms_unadj |> glimpse()


```

### Demography adjusted Models
```{r}

deprv_dms_demo_adj_cxph <-
  survival::coxph(formula = Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_dms,
                                    data    = ehr_cpr_cens11_cohrt,
                                    id      = c_alf_pe)
summary(deprv_dms_demo_adj_cxph)

# compute population AF: set each dimension to none
tic()
Sys.time()
deprv_dms_demo_adj_to_none <-
  paf_deprv_no_to_zero <-
    calc_deprv_level_paf(
      ftd_mdl = deprv_dms_demo_adj_cxph,
      data = ehr_cpr_cens11_cohrt,
      fct_var = m_hh_deprv_dms,
      ref_lvl = "None",
      boot_rep = 100,
      t_vector = follow_up_yr
    )
Sys.time()
toc() #715.32 sec elapsed
deprv_dms_demo_adj_to_none |> glimpse()

# compute population AF: remove each unitary dimension in turn
tic()
Sys.time()
paf_dms_demo_adj <- calc_dprv_dms_rmvl_paf(
  ftd_mdl = deprv_dms_demo_adj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_dms,
  t_vector = seq.int(0, 11),
  output_file = "paf_dms_demo_adj.csv"
)
Sys.time()
toc() #1625.84 sec elapsed (30 mins)
paf_dms_demo_adj |> glimpse()

```

### Full model
```{r}

names(cx_mdls_multi_full)

# Model full_2b: 
# Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + 
#     m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_dms + 
#     c_cgm + c_ptb + c_bw + m_alc + m_lrn + m_mnt + m_sbm + m_smk

deprv_dms_full_adj_cxph <-
  survival::coxph(formula = Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_dms + c_cgm + c_ptb + c_bw + m_alc + m_lrn + m_mnt + m_sbm + m_smk,
                                    data    = ehr_cpr_cens11_cohrt,
                                    id      = c_alf_pe)
summary(deprv_dms_full_adj_cxph)

# compute population AF: set each dimension to none
tic()
Sys.time()
deprv_dms_full_adj_to_none <-
  paf_deprv_no_to_zero <-
    calc_deprv_level_paf(
      ftd_mdl = deprv_dms_full_adj_cxph,
      data = ehr_cpr_cens11_cohrt,
      fct_var = m_hh_deprv_dms,
      ref_lvl = "None",
      boot_rep = 100,
      t_vector = follow_up_yr
    )
Sys.time()
toc() #944.53 sec elapsed
deprv_dms_full_adj_to_none |> glimpse()

# compute population AF: remove each unitary dimension in turn
tic()
Sys.time()
paf_dms_full_adj <- calc_dprv_dms_rmvl_paf(
  ftd_mdl = deprv_dms_full_adj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_dms,
  t_vector = seq.int(0, 11),
  output_file = "paf_dms_full_adj.csv"
)
Sys.time()
toc() #
paf_dms_full_adj |> glimpse()

```

### PAF Excel export
```{r}
# Moving deprivation domains to none
dprv_dms_2_non_tbl <- 
  bind_rows(list("Unadjusted" = deprv_dms_to_none,
               "Demography-Adjusted" = deprv_dms_demo_adj_to_none,
               "Fully-Adjusted" = deprv_dms_full_adj_to_none),
          .id = "adj_lvl") |> 
  glimpse()

# remove each unitary dimension in turn
dprv_dms_rmvl_tbl <-
  bind_rows(list("Unadjusted" = paf_dms_unadj,
               "Demography-Adjusted" = paf_dms_demo_adj,
               "Fully-Adjusted" = paf_dms_full_adj),
          .id = "adj_lvl") |> 
  glimpse()

# save list of named dataframes as xlsx with one df per sheet
    writexl::write_xlsx(list("dprv_dms_2_non" = dprv_dms_2_non_tbl,
                             "dprv_dms_rmvl" = dprv_dms_rmvl_tbl),
                        path = "outputs\\paf_dprv_dms.xlsx")

```

### Plots: PAF deprivation domains
```{r}

# create plots
## Plot counterfactual removal of each unitary deprivation dimension
paf_dms_rmvl_plot <- plot_dprv_dms_rmvl_paf(
  paf_results = list("Unadjusted" = paf_dms_unadj,
                     "Demography-Adjusted" = paf_dms_demo_adj,
                     "Fully-Adjusted" = paf_dms_full_adj
                     ),
  time_point = 6
)
# paf_dms_rmvl_plot

## Plot counterfactual move of each dms combination to none
paf_dms_to_none_plot <- plot_deprv_level_paf(
  paf_results = list("Unadjusted" = deprv_dms_to_none,
                     "Demography-Adjusted" = deprv_dms_demo_adj_to_none,
                     "Fully-Adjusted" = deprv_dms_full_adj_to_none
                     ),
  time_point = 6
)
# paf_dms_to_none_plot

# mosaic the deprivation domain PAF plots
# unit(c(0,30,0,0) = top, right, bottom, left
paf_dms_all_stch_plts <-
  ((paf_dms_to_none_plot + theme(plot.margin = unit(c(0,0,0.5,0), "in")))/paf_dms_rmvl_plot) +
    plot_layout(guides = 'collect',
                axis_titles = "collect",
                heights = c(1,3)) &
    theme(legend.position='bottom')

ggsave(
  filename = "outputs\\paf_dms_all_stch_plts.png",
  plot = paf_dms_all_stch_plts,
  # scale = 0.75,
  height = 15,
  width = 14,
  # units = 'cm',
  # dpi = 900
)

# Extract PAF tables

calc_deprv_level_paf(deprv_dms_to_none)

gc()
save.image()
```

## Deprivation counts
Consider the following counterfactual scenarios (and their associated plots)
* Move each count of deprivation to zero (eradicating HH deprivation altogether). 
* Reduce counts of HH deprivation (a) by 1, and (b) by 2
* reduce deprivation in HHs with (a) counts above 1 to 1 & (b) counts above 2 to 2


### Unadjusted (bivariate) Models
```{r}

#     fit bivariate survival model
deprv_no_unadj_cxph <-
  survival::coxph(formula = Surv(c_age_1st_evnt, c_frst_cpr) ~ m_hh_deprv_no,
                                    data    = ehr_cpr_cens11_cohrt,
                                    id      = c_alf_pe)

summary(deprv_no_unadj_cxph)

# age vector of interest
follow_up_yr <- seq.int(0, 11)
follow_up_yr

# compute population AF: set each deprivation count to 0 in turns
tic()
Sys.time()
paf_deprv_no_to_0_unadj <-
  calc_deprv_level_paf(
    ftd_mdl = deprv_no_unadj_cxph,
    data = ehr_cpr_cens11_cohrt,
    fct_var = m_hh_deprv_no,
    ref_lvl = "0",
    boot_rep = 100,
    t_vector = follow_up_yr
  )
Sys.time()
toc() #112.39 sec elapsed

paf_deprv_no_to_0_unadj |> 
  filter(Time == 6) |> 
  view("unadj")

# compute population AF: reduce deprivation by 1
tic()
Sys.time()
paf_deprv_rdtn_by1_unadj <-
  calc_dprv_no_rdctn_paf(
  ftd_mdl = deprv_no_unadj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_no,
  dprv_rdctn_no = 1,
  t_vector = follow_up_yr,
  output_file = "paf_deprv_rdtn_by1_unadj.csv"
)
Sys.time()
toc() #123.25 sec elapsed

paf_deprv_rdtn_by1_unadj |> glimpse()

# compute population AF: reduce deprivation by 2
tic()
Sys.time()
paf_deprv_rdtn_by2_unadj <-
  calc_dprv_no_rdctn_paf(
  ftd_mdl = deprv_no_unadj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_no,
  dprv_rdctn_no = 2,
  t_vector = follow_up_yr,
  output_file = "paf_deprv_rdtn_by2_unadj.csv"
)
Sys.time()
toc() #110.7 sec elapsed

paf_deprv_rdtn_by2_unadj |> glimpse()

# # move only deprivation counts above 1 to level 1.
# tic()
# Sys.time()
# paf_deprv_gt1_to_1_unadj <-
#   calc_deprv_level_paf(
#     ftd_mdl = deprv_no_unadj_cxph,
#     data = ehr_cpr_cens11_cohrt,
#     fct_var = m_hh_deprv_no,
#     ref_lvl = "1",
#     excld_lvl = "0",
#     boot_rep = 100,
#     t_vector = follow_up_yr
#   )
# Sys.time()
# toc() #108.61 sec elapsed
# 
# paf_deprv_gt1_to_1_unadj |> glimpse()
```

### Demography adjusted models
```{r}

covar_all

# refit coxph model
frml_a <- as.formula(paste0("Surv(c_age_1st_evnt, c_frst_cpr) ~ ", covar_all[[14]]))
deparse1(frml_a)
# cat("\nModel", i, "\n")
# print(frml_a)

# extract formular from originally fitted model
formula(deparse1(cx_mdl_multi[["demo2a"]]$formula))

#     fit demography adjusted survival model
deprv_no_demo_adj_cxph <-
  survival::coxph(formula = Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_no,
                                    data    = ehr_cpr_cens11_cohrt,
                                    id      = c_alf_pe)

summary(deprv_no_demo_adj_cxph)

# compute population AF: set deprivation to 0
tic()
Sys.time()
paf_deprv_no_to_0_demo_adj <-
  calc_deprv_level_paf(
    ftd_mdl = deprv_no_demo_adj_cxph,
    data = ehr_cpr_cens11_cohrt,
    fct_var = m_hh_deprv_no,
    ref_lvl = "0",
    boot_rep = 100,
    t_vector = follow_up_yr
  )
Sys.time()
toc() #

paf_deprv_no_to_0_demo_adj |> glimpse()
paf_deprv_no_to_0_demo_adj |> 
  filter(Time == 6) |> 
  view("demo-adj")

# compute population AF: reduce deprivation by 1
tic()
Sys.time()
paf_deprv_rdtn_by1_demo_adj <- calc_dprv_no_rdctn_paf(
  ftd_mdl = deprv_no_demo_adj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_no,
  dprv_rdctn_no = 1,
  t_vector = follow_up_yr,
  output_file = "paf_deprv_rdtn_by1_demo_adj.csv"
)
Sys.time()
toc() #

paf_deprv_rdtn_by1_demo_adj |> glimpse()

# compute population AF: reduce deprivation by 2
tic()
Sys.time()
paf_deprv_rdtn_by2_demo_adj <-
  calc_dprv_no_rdctn_paf(
  ftd_mdl = deprv_no_demo_adj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_no,
  dprv_rdctn_no = 2,
  t_vector = follow_up_yr,
  output_file = "paf_deprv_rdtn_by2_demo_adj.csv"
)
Sys.time()
toc() #

paf_deprv_rdtn_by2_demo_adj |> glimpse()

# # move only deprivation counts above 1 to level 1.
# tic()
# Sys.time()
# paf_deprv_gt1_to_1_demo_adj <-
#   calc_deprv_level_paf(
#     ftd_mdl = deprv_no_demo_adj_cxph,
#     data = ehr_cpr_cens11_cohrt,
#     fct_var = m_hh_deprv_no,
#     ref_lvl = "1",
#     excld_lvl = "0",
#     boot_rep = 100,
#     t_vector = follow_up_yr
#   )
# Sys.time()
# toc()
# 
# paf_deprv_gt1_to_1_demo_adj |> glimpse()

gc()
# save.image()
```

### Full model
```{r}

names(cx_mdls_multi_full)

# Model full_2a: 
# Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + 
#     m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_no + 
#     c_cgm + c_ptb + c_bw + m_alc + m_lrn + m_mnt + m_sbm + m_smk

#     fit demography adjusted survival model
deprv_no_full_adj_cxph <-
  survival::coxph(formula = Surv(c_age_1st_evnt, c_frst_cpr) ~ c_sex + c_ethn + m_age_cat + m_edu + m_hh_child_depndt + m_hh_adult + m_mar_status + m_hh_deprv_no + c_cgm + c_ptb + c_bw + m_alc + m_lrn + m_mnt + m_sbm + m_smk,
                                    data    = ehr_cpr_cens11_cohrt,
                                    id      = c_alf_pe)

summary(deprv_no_full_adj_cxph)

# compute population AF: set deprivation to 0
tic()
Sys.time()
paf_deprv_no_to_0_full_adj <-
  calc_deprv_level_paf(
    ftd_mdl = deprv_no_full_adj_cxph,
    data = ehr_cpr_cens11_cohrt,
    fct_var = m_hh_deprv_no,
    ref_lvl = "0",
    boot_rep = 100,
    t_vector = follow_up_yr
  )
Sys.time()
toc() #207.26 sec elapsed

paf_deprv_no_to_0_full_adj |> glimpse()
paf_deprv_no_to_0_full_adj |> 
  filter(Time == 6) |> 
  view("full-adj")

# compute population AF: reduce deprivation by 1
tic()
Sys.time()
paf_deprv_rdtn_by1_full_adj <- calc_dprv_no_rdctn_paf(
  ftd_mdl = deprv_no_full_adj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_no,
  dprv_rdctn_no = 1,
  t_vector = follow_up_yr,
  output_file = "paf_deprv_rdtn_by1_full_adj.csv"
)
Sys.time()
toc() #183.8 sec elapsed

paf_deprv_rdtn_by1_full_adj |> glimpse()

# compute population AF: reduce deprivation by 2
tic()
Sys.time()
paf_deprv_rdtn_by2_full_adj <-
  calc_dprv_no_rdctn_paf(
  ftd_mdl = deprv_no_full_adj_cxph,
  data = ehr_cpr_cens11_cohrt,
  fct_var = m_hh_deprv_no,
  dprv_rdctn_no = 2,
  t_vector = follow_up_yr,
  output_file = "paf_deprv_rdtn_by2_full_adj.csv"
)
Sys.time()
toc() #207.76 sec elapsed

paf_deprv_rdtn_by2_full_adj |> glimpse()

# # move only deprivation counts above 1 to level 1.
# tic()
# Sys.time()
# paf_deprv_gt1_to_1_full_adj <-
#   calc_deprv_level_paf(
#     ftd_mdl = deprv_no_full_adj_cxph,
#     data = ehr_cpr_cens11_cohrt,
#     fct_var = m_hh_deprv_no,
#     ref_lvl = "1",
#     excld_lvl = "0",
#     boot_rep = 100,
#     t_vector = follow_up_yr
#   )
# Sys.time()
# toc() #214.85 sec elapsed
# 
# paf_deprv_gt1_to_1_full_adj |> glimpse()

gc()

```

### PAF count dprv excel export
```{r}
# Moving deprivation count to 0
dprv_2_non_tbl <- 
  bind_rows(list("Unadjusted" = paf_deprv_no_to_0_unadj,
               "Demography-Adjusted" = paf_deprv_no_to_0_demo_adj,
               "Fully-Adjusted" = paf_deprv_no_to_0_full_adj),
          .id = "adj_lvl") |> 
  glimpse()

# reduce deprivation count by 1
dprv_rdtn_by1_tbl <-
  bind_rows(list("Unadjusted" = paf_deprv_rdtn_by1_unadj,
               "Demography-Adjusted" = paf_deprv_rdtn_by1_demo_adj,
               "Fully-Adjusted" = paf_deprv_rdtn_by1_full_adj),
          .id = "adj_lvl") |> 
  glimpse()

# reduce deprivation count by 2
dprv_rdtn_by2_tbl <-
  bind_rows(list("Unadjusted" = paf_deprv_rdtn_by2_unadj,
               "Demography-Adjusted" = paf_deprv_rdtn_by2_demo_adj,
               "Fully-Adjusted" = paf_deprv_rdtn_by2_full_adj),
          .id = "adj_lvl") |> 
  glimpse()

# save list of named dataframes as xlsx with one df per sheet
    writexl::write_xlsx(list("dprv_no_2_non" = dprv_2_non_tbl,
                             "dprv_rdtn_by1" = dprv_rdtn_by1_tbl,
                             "dprv_rdtn_by2" = dprv_rdtn_by2_tbl),
                        path = "outputs\\paf_dprv_count.xlsx")
```

### Plots: PAF deprivation domains
```{r}

# writexl::write_xlsx(cx_mdl_full_af_all, path =
#                           "outputs\\cx_mdl_full_af_tbls_all.xlsx")

# create plots
## Plot population AF: reduce all deprivation counts to 0
paf_plot_deprv_no_to_0 <- plot_dprv_no_rdctn_paf(
  paf_results = list(
    "Unadjusted" = paf_deprv_no_to_0_unadj,
    "Demography adjusted" = paf_deprv_no_to_0_demo_adj,
    "Fully adjusted" = paf_deprv_no_to_0_full_adj
  ),
  output_stem = "paf_dprvtn_no_to_0",
  title = "AF after reducing household deprivation counts to Zero",
  time_point = 6
)
# paf_plot_deprv_no_to_0

## Plot population AF: reduce deprivation by 1
paf_plot_deprv_rdtn_by1 <- 
  plot_dprv_no_rdctn_paf(
  paf_results = list(
    "Unadjusted" = paf_deprv_rdtn_by1_unadj,
    "Demography adjusted" = paf_deprv_rdtn_by1_demo_adj,
    "Fully adjusted" = paf_deprv_rdtn_by1_full_adj
  ),
  output_stem = "paf_dprvtn_cnt_rdtn_by1",
  title = "AF after reducing household deprivation counts by 1",
  time_point = 6
)
# paf_plot_deprv_rdtn_by1

## Plot population AF: reduce deprivation by 2
paf_plot_deprv_rdtn_by2 <- 
  plot_dprv_no_rdctn_paf(
  paf_results = list(
    "Unadjusted" = paf_deprv_rdtn_by2_unadj,
    "Demography adjusted" = paf_deprv_rdtn_by2_demo_adj,
    "Fully adjusted" = paf_deprv_rdtn_by2_full_adj
  ),
  output_stem = "paf_dprvtn_cnt_rdtn_by2",
  title = "AF after reducing household deprivation counts by 2",
  time_point = 6
)
# paf_plot_deprv_rdtn_by2

# mosaic the deprivation count PAF plots
paf_pdrv_no_all_stch_plts <-
  (paf_plot_deprv_no_to_0 | paf_plot_deprv_rdtn_by1 | paf_plot_deprv_rdtn_by2)  +
    plot_layout(guides = 'collect',
                axes = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

ggsave(
  filename = "outputs\\paf_pdrv_no_all_stch_plts.png",
  plot = paf_pdrv_no_all_stch_plts,
  # scale = 0.75,
  height = 6,
  width = 18,
  # units = 'cm',
  # dpi = 900
)  

gc()
save.image()
```

# Adjusted Survival curves
## Demographic Models
```{r}
names(cx_mdls_multi_vars)

# print only formula to fit for documentation
for (i in seq_along(cx_mdls_multi_vars)) {
  if (i == 1) {
    outa_frml_no <- length(cx_mdls_multi_vars)
    cat("\nOuter formula (Total:", outa_frml_no, ")\n")
    
    cxph_adj_demo_tbls_all <- list()
    
    #     ensure proper dataframe
    ehr_cpr_cens11_cohrt_df <- as.data.frame(ehr_cpr_cens11_cohrt)
  }
  
  #   prep formula, fit models, process & output coefficients
  outa_frml <- paste0("Surv(c_age_1st_evnt, c_frst_cpr) ~ ",
                      paste(cx_mdls_multi_vars[[i]], collapse = " + "))
  
  cat("\n----------------------------\n")
  mdl_nme <- names(cx_mdls_multi_vars[i])
  cat("Model Name:", mdl_nme)
  cat("\n", i, as.formula(outa_frml) %>% deparse1())
  
  # names(cx_mdls_multi_vars)
  
  #   control for each factor as a grouping variable in the model
  for (j in seq_along(cx_mdls_multi_vars[[i]])) {
    if (i == 1 & j == 1) {
      inner_frml_no <- length(cx_mdls_multi_vars[[i]])
      cat("\n\tInner formula (Total:", inner_frml_no, ")\n")
      
      cxph_adj_demo_mdls <- list()
      cxph_adj_demo_plts <- list()
      cxph_adj_demo_tbls <- list()
    }
    
# cx_mdls_multi_vars[[1]]
    
    #     stratifying variables (for adjusted model)
    strata_var <- cx_mdls_multi_vars[[i]][j]
    #     name model of strata
    adj_mdl_nme <- paste0(mdl_nme, ".", strata_var)
    #     upgrade formula with strata
    frml_strata <- paste0(outa_frml, " + strata(", strata_var, ")")
    cat("\n\t", j, as.formula(frml_strata) %>% deparse1(), "\n")
    
    try(# using coxph + ggsurvfit package
      #   create model
      cxph_adj_demo_mdls[[adj_mdl_nme]] <-
        coxph(formula = as.formula(frml_strata), data = ehr_cpr_cens11_cohrt_df) %>%
        ggsurvfit::survfit2())
    
    # NO LONGER NEEDED SINCE CAN'T PASS SAIL'S SDC
    # if ("survfit2" %in% class(cxph_adj_demo_mdls[[adj_mdl_nme]])) {
    #   #   tidy tables of the model, for flexible plotting elsewhere
    #   cxph_adj_demo_tbls[[adj_mdl_nme]] <-
    #     cxph_adj_demo_mdls[[adj_mdl_nme]] %>%
    #     tidy_survfit()
    #   # demo_mdl_adj_tbl
    # }
    
    # class(cxph_adj_demo_mdls[[1]])
    
    #     proper formating of print name
    # list all potantial exposures/covariates
    if (strata_var == "c_sex") {
      strata_nme = "Child Sex"
    }
    if (strata_var == "c_ethn") {
      strata_nme = "Child ethnicity"
    }
    if (strata_var == "c_cgm") {
      strata_nme = "Child congenital anomalies"
    }
    if (strata_var == "c_ptb") {
      strata_nme = "Child preterm birth"
    }
    if (strata_var == "c_bw") {
      strata_nme = "Child birth weight"
    }
    if (strata_var == "m_age_cat") {
      strata_nme = "Mother age at birth"
    }
    if (strata_var == "m_ethnic") {
      strata_nme = "Mother ethnicity"
    }
    if (strata_var == "m_edu") {
      strata_nme = "Mother eductation"
    }
    if (strata_var == "m_hh_child_depndt") {
      strata_nme = "Mother HH dependent children"
    }
    if (strata_var == "m_hh_adult") {
      strata_nme = "Mother HH adults"
    }
    if (strata_var == "m_mar_status") {
      strata_nme = "Mother marital status"
    }
    if (strata_var == "m_fam_status") {
      strata_nme = "Mother family status"
    }
    if (strata_var == "m_uk_stay_len") {
      strata_nme = "Mother length of UK stay"
    }
    if (strata_var == "m_hh_deprv_no") {
      strata_nme = "Mother HH deprivation count"
    }
    if (strata_var == "m_hh_deprv_dms") {
      strata_nme = "Mother HH deprivation dimensions"
    }
    if (strata_var == "m_twns2011") {
      strata_nme = "Mother Townsend 2011 quintile"
    }
    if (strata_var == "m_alc") {
      strata_nme = "Mother alcohol problems"
    }
    if (strata_var == "m_lrn") {
      strata_nme = "Mother learning difficulties/disabilities"
    }
    if (strata_var == "m_mnt") {
      strata_nme = "Mother mental health issues"
    }
    if (strata_var == "m_sbm") {
      strata_nme = "Mother substance use"
    }
    if (strata_var == "m_smk") {
      strata_nme = "Mother smoking"
    }
    
    if ("survfit2" %in% class(cxph_adj_demo_mdls[[adj_mdl_nme]])) {
      #   create plots from models
      cxph_adj_demo_plts[[adj_mdl_nme]] <-
        plot_sdc_smooth_survfit(
          survfit_obj = cxph_adj_demo_mdls[[adj_mdl_nme]],
          type = "risk",
          method = "loess",
          span = 0.15, #reduce to reduce smoothing
          smooth_n = 25, #tends to affect the CIs
          add_noise = T,
          noise_scope = "tail",
          noise_tail_n = 3,
          noise_sd = 0.00005,
          show_ci = T,
          subtitle = strata_nme,
          caption = paste0("Demography Adjusted Model (", mdl_nme, ")")
        ) +
        ggplot2::scale_x_continuous(#limits = c(0,11.5),
                                    breaks = seq.int(0,10,2),
                                    expand = expansion(add = 0.1)
                                    ) #+
        # ggplot2::scale_y_continuous(labels = scales::label_percent(accuracy = 0.01),
        #                         expand = expansion(mult = 0.002)
        #                         )
      
      ?scales::label_percent
      
      ## print all strata
      # cx_mdls_multi_vars
      
      #   adjust plot's y coordinates of relevant exposures (by zooming)
      if (strata_var == "m_mar_status") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.015))
      }
      if (strata_var == "m_age_cat") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.004))
      }
      if (strata_var == "m_edu") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.0085))#
      }
      if (strata_var == "c_sex") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.00175))#
      }
      if (strata_var == "m_hh_adult") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.003))
      }
      if (strata_var == "m_hh_child_depndt") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.003))
      }
      if (strata_var == "c_ethn") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.002))
      }
      if (strata_var == "m_hh_deprv_no") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.01)) #8
      }
      if (strata_var == "m_hh_deprv_dms") {
        cxph_adj_demo_plts[[adj_mdl_nme]] <-
          cxph_adj_demo_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.0085)) #8
      }

        #   plots needing wider width
      if (strata_var %in% c("m_fam_status", "m_mar_status")) {
        wdth = 20
      } else if (strata_var %in% c("m_hh_deprv_dms", "m_ethnic")) {
        wdth = 25
      }
      else{
        wdth = 15
      }
  
      #make outputs directory (if it doesn't already exist)
      plt_dir <- paste0("outputs\\demo_adj_plts_sdc\\demo_adj_plts_", mdl_nme)
      # cat("\n\tDirectory:", plt_dir, "\n")
      
      if (!dir.exists(plt_dir)) {
        dir.create(plt_dir)
      }
      
      ggsave(
        filename = paste0(plt_dir, "\\", mdl_nme, "_", strata_var, ".png"),
        plot = cxph_adj_demo_plts[[adj_mdl_nme]],
        scale = 0.9,
        height = 20,
        width = wdth,
        units = 'cm',
        dpi = 600
      )
    }
    
    # NO LONGER NEEDED SINCE CAN'T PASS SAIL'S SDC
    # if (j == inner_frml_no) {
    #   cxph_adj_demo_tbls_all[[mdl_nme]] <-
    #     bind_rows(cxph_adj_demo_tbls, .id = "model_name") %>%
    #     select(-strata_label)
    # }
    
  }
  
  if (i == outa_frml_no) {
        # NO LONGER NEEDED SINCE CAN'T PASS SAIL'S SDC
      #   save list of named df to excel spread sheet eith named tabs
      # writexl::write_xlsx(cxph_adj_demo_tbls_all, path = "outputs\\cxph_adj_demo_tbls_all.xlsx")
      
      cat("Cleaning & saving workspace")
      #     clean and save workspace
      rm(
        plt_dir,
        fct_no,
        strata_nme,
        strata_var,
        frml_strata,
        adj_mdl_nme,
        outa_frml,
        outa_frml_no,
        mdl_nme,
        mdl_cnt,
        inner_frml_no
      )
      gc()
      # save.image()
    }
}
  

names(cxph_adj_demo_plts)
```

## Full Models
```{r}

# replace cx_mdls_multi_vars with cx_mdls_multi_vars
seq_along(cx_mdls_multi_full)

# print only formula to fit for documentation
for (i in seq_along(cx_mdls_multi_full)) {
  if (i == 1) {
    outa_frml_no <- length(cx_mdls_multi_full)
    cat("\nOuter formula (Total:", outa_frml_no, ")\n")
    
    cxph_adj_full_tbls_all <- list()
    
    #     ensure proper dataframe
    ehr_cpr_cens11_cohrt_df <- as.data.frame(ehr_cpr_cens11_cohrt)
  }
  
  #   prep formula, fit models, process & output coefficients
  outa_frml <- paste0("Surv(c_age_1st_evnt, c_frst_cpr) ~ ",
                      paste(cx_mdls_multi_full[[i]], collapse = " + "))
  
  cat("\n----------------------------\n")
  mdl_nme <- names(cx_mdls_multi_full[i])
  cat("Model Name:", mdl_nme)
  cat("\n", i, as.formula(outa_frml) %>% deparse1())
  
  #   control for each factor as a grouping variable in the model
  for (j in seq_along(cx_mdls_multi_full[[i]])) {
    if (i == 1 & j == 1) {
      inner_frml_no <- length(cx_mdls_multi_full[[i]])
      cat("\n\tInner formula (Total:", inner_frml_no, ")\n")
      
      cxph_adj_full_mdls <- list()
      cxph_adj_full_plts <- list()
      cxph_adj_full_tbls <- list()
    }
    
    #     stratifying variables (for adjusted model)
    strata_var <- cx_mdls_multi_full[[i]][j]
    #     name model of strata
    adj_mdl_nme <- paste0(mdl_nme, ".", strata_var)
    #     upgrade formula with strata
    frml_strata <- paste0(outa_frml, " + strata(", strata_var, ")")
    cat("\n\t", j, as.formula(frml_strata) %>% deparse1(), "\n")
    
    try(# using coxph + ggsurvfit package
      #   create model
      cxph_adj_full_mdls[[adj_mdl_nme]] <-
        coxph(formula = as.formula(frml_strata), data = ehr_cpr_cens11_cohrt_df) %>%
        ggsurvfit::survfit2())
    
    # NO LONGER NEEDED SINCE CAN'T PASS SAIL'S SDC
    # if ("survfit2" %in% class(cxph_adj_full_mdls[[adj_mdl_nme]])) {
    #   #   tidy tables of the model, for flexible plotting elsewhere
    #   cxph_adj_full_tbls[[adj_mdl_nme]] <-
    #     cxph_adj_full_mdls[[adj_mdl_nme]] %>%
    #     tidy_survfit()
    #   # full_mdl_adj_tbl
    # }
    
    # class(cxph_adj_full_mdls[[1]])
    
    #     proper formating of print name
    # list all potantial exposures/covariates
    if (strata_var == "c_sex") {
      strata_nme = "Child Sex"
    }
    if (strata_var == "c_ethn") {
      strata_nme = "Child ethnicity"
    }
    if (strata_var == "c_cgm") {
      strata_nme = "Child congenital anomalies"
    }
    if (strata_var == "c_ptb") {
      strata_nme = "Child preterm birth"
    }
    if (strata_var == "c_bw") {
      strata_nme = "Child birth weight"
    }
    if (strata_var == "m_age_cat") {
      strata_nme = "Mother age at birth"
    }
    if (strata_var == "m_ethnic") {
      strata_nme = "Mother ethnicity"
    }
    if (strata_var == "m_edu") {
      strata_nme = "Mother eductation"
    }
    if (strata_var == "m_hh_child_depndt") {
      strata_nme = "Mother HH dependent children"
    }
    if (strata_var == "m_hh_adult") {
      strata_nme = "Mother HH adults"
    }
    if (strata_var == "m_mar_status") {
      strata_nme = "Mother marital status"
    }
    if (strata_var == "m_fam_status") {
      strata_nme = "Mother family status"
    }
    if (strata_var == "m_uk_stay_len") {
      strata_nme = "Mother length of UK stay"
    }
    if (strata_var == "m_hh_deprv_no") {
      strata_nme = "Mother HH deprivation count"
    }
    if (strata_var == "m_hh_deprv_dms") {
      strata_nme = "Mother HH deprivation dimensions"
    }
    if (strata_var == "m_twns2011") {
      strata_nme = "Mother Townsend 2011 quintile"
    }
    if (strata_var == "m_alc") {
      strata_nme = "Mother alcohol problems"
    }
    if (strata_var == "m_lrn") {
      strata_nme = "Mother learning difficulties/disabilities"
    }
    if (strata_var == "m_mnt") {
      strata_nme = "Mother mental health issues"
    }
    if (strata_var == "m_sbm") {
      strata_nme = "Mother substance use"
    }
    if (strata_var == "m_smk") {
      strata_nme = "Mother smoking"
    }
    
    if ("survfit2" %in% class(cxph_adj_full_mdls[[adj_mdl_nme]])) {
      #   create plots from models
      cxph_adj_full_plts[[adj_mdl_nme]] <-
        plot_sdc_smooth_survfit(
          survfit_obj = cxph_adj_full_mdls[[adj_mdl_nme]],
          type = "risk",
          method = "loess",
          span = 0.15, #reduce to reduce smoothing
          smooth_n = 25, #tends to affect the CIs
          add_noise = T,
          noise_scope = "tail",
          noise_tail_n = 3,
          noise_sd = 0.00005,
          show_ci = T,
          subtitle = strata_nme,
          caption = paste0("Fully Adjusted Model (", mdl_nme, ")")
        ) +
        ggplot2::scale_x_continuous(#limits = c(0,11.5),
                                    breaks = seq.int(0,10,2),
                                    expand = expansion(add = 0.1)
                                    )
      
      #   adjust plot's y coordinates of relevant exposures (by zooming)
      if (strata_var == "m_mar_status") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.015))
      }
      if (strata_var == "m_age_cat") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.004))
      }
      if (strata_var == "m_edu") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.0085))#
      }
      if (strata_var == "c_sex") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.00175))#
      }
      if (strata_var == "m_hh_adult") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.003))
      }
      if (strata_var == "m_hh_child_depndt") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.003))
      }
      if (strata_var == "c_ethn") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.002))
      }
      if (strata_var == "m_hh_deprv_no") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.01)) #8
      }
      if (strata_var == "m_hh_deprv_dms") {
        cxph_adj_full_plts[[adj_mdl_nme]] <-
          cxph_adj_full_plts[[adj_mdl_nme]] + coord_cartesian(ylim = c(0, 0.0085)) #8
      }
      
      
      #   plots needing wider width
      if (strata_var %in% c("m_fam_status", "m_mar_status")) {
        wdth = 20
      } else if (strata_var %in% c("m_hh_deprv_dms", "m_ethnic")) {
        wdth = 25
      }
      else{
        wdth = 15
      }
      
      #make outputs directory (if it doesn't already exist)
      plt_dir <- paste0("outputs\\full_adj_plts_sdc\\full_adj_plts_", mdl_nme)
      # cat("\n\tDirectory:", plt_dir, "\n")
      
      if (!dir.exists(plt_dir)) {
        dir.create(plt_dir)
      }
      
      ggsave(
        filename = paste0(plt_dir, "\\", mdl_nme, "_", strata_var, ".png"),
        plot = cxph_adj_full_plts[[adj_mdl_nme]],
        scale = 0.9,
        height = 20,
        width = wdth,
        units = 'cm',
        dpi = 600
      )
    }
    
    # NO LONGER NEEDED SINCE CAN'T PASS SAIL'S SDC
    # if (j == inner_frml_no) {
    #   cxph_adj_full_tbls_all[[mdl_nme]] <-
    #     bind_rows(cxph_adj_full_tbls, .id = "model_name") %>%
    #     select(-strata_label)
    # }
    
  }
  
  if (i == outa_frml_no) {
    # NO LONGER NEEDED SINCE CAN'T PASS SAIL'S SDC
    #   save list of named df to excel spread sheet eith named tabs
    # writexl::write_xlsx(cxph_adj_full_tbls_all, path = "outputs\\full_adj_plts_sdc\\cxph_adj_full_tbls_all_sdc.xlsx")
    
    cat("Cleaning & saving workspace")
    #     clean and save workspace
    rm(
      plt_dir,
      fct_no,
      strata_nme,
      strata_var,
      frml_strata,
      adj_mdl_nme,
      outa_frml,
      outa_frml_no,
      mdl_nme,
      mdl_cnt,
      inner_frml_no
    )
    gc()
    # save.image()
  }
}

cxph_adj_full_tbls_all[[1]] %>%
  glimpse

```

## Join Survival curves: unadj, demo adj, full adj
### Deprivation Domains
```{r}

# unadj
names(all_klpn_plts)

# demo adj models
names(cxph_adj_demo_plts)

# full models
names(cxph_adj_full_plts)

# stitch_hh_dprv_no
m_hh_deprv_no_stp_adj <- 
  ((all_klpn_plts[["m_hh_deprv_no"]] + labs(
    title = "Count of household deprivation dimensions",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2a.m_hh_deprv_no"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2a.m_hh_deprv_no"]] + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  ))
  ) +
    plot_annotation(title = "Count of HH deprivation dimensions",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\m_hh_deprv_no_stp_adj.png",
#   plot = m_hh_deprv_no_stp_adj,
#   scale = 0.75,
#   height = 20,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# )

# stitch_hh_dprv_combinations
m_hh_deprv_dms_stp_adj <- 
  ((all_klpn_plts[["m_hh_deprv_dms"]] + labs(
    title = "Combinations of household deprivation dimensions",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2b.m_hh_deprv_dms"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2b.m_hh_deprv_dms"]]) + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  )) +
    plot_annotation(title = "Combinations of HH deprivation dimensions",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\m_hh_deprv_dms_stp_adj.png",
#   plot = m_hh_deprv_dms_stp_adj,
#   scale = 0.75,
#   height = 25,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# ) 

# ?ggsave

# save full risk curves
surv_hh_deprv_all_stp_adj <-
  m_hh_deprv_no_stp_adj / m_hh_deprv_dms_stp_adj

ggsave(
  filename = "outputs\\surv_hh_deprv_all_stp_adj.png",
  plot = surv_hh_deprv_all_stp_adj,
  scale = 0.6,
  height = 50,
  width = 45,
  units = 'cm',
  dpi = 900
)

# save.image()
```

### Child's demography
c_sex + c_ethn + m_age_cat + m_edu + m_mar_status + m_hh_child_depndt + m_hh_adult
```{r}
# risk curves for other demographic variables
#Child's Sex
surv_c_sex_stp_adj <- 
  ((all_klpn_plts[["c_sex"]] + labs(
    title = "Child Sex",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2a.c_sex"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2a.c_sex"]] + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  ))
  ) +
    plot_annotation(title = "Child Sex",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\surv_c_sex_stp_adj.png",
#   plot = surv_c_sex_stp_adj,
#   scale = 0.7,
#   height = 20,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# )

#Child ethnicity
surv_c_ethn_stp_adj <- 
  ((all_klpn_plts[["c_ethn"]] + labs(
    title = "Child Ethnicity",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2a.c_ethn"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2a.c_ethn"]] + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  ))
  ) +
    plot_annotation(title = "Child Ethnicity",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\surv_c_ethn_stp_adj.png",
#   plot = surv_c_ethn_stp_adj,
#   scale = 0.7,
#   height = 20,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# )

# All Child demography
surv_c_demo_all_stp_adj <- (surv_c_sex_stp_adj / surv_c_ethn_stp_adj) +
    plot_layout(axis_titles = 'collect')

ggsave(
  filename = "outputs\\surv_c_demo_all_stp_adj.png",
  plot = surv_c_demo_all_stp_adj,
  scale = 0.6,
  height = 45,
  width = 45,
  units = 'cm',
  dpi = 900
)

gc()
# save.image()
```

### Mother's Demography
```{r}
#Mother's Age
surv_m_age_stp_adj <- 
  ((all_klpn_plts[["m_age_cat"]] + labs(
    title = "Mother's Age",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2a.m_age_cat"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2a.m_age_cat"]] + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  ))
  ) +
    plot_annotation(title = "Mother's Age",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\surv_m_age_stp_adj.png",
#   plot = surv_m_age_stp_adj,
#   scale = 0.7,
#   height = 20,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# )

#Mother's Education
surv_m_edu_stp_adj <- 
  ((all_klpn_plts[["m_edu"]] + labs(
    title = "Mother's Education",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2a.m_edu"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2a.m_edu"]] + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  ))
  ) +
    plot_annotation(title = "Mother's Education",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\surv_m_edu_stp_adj.png",
#   plot = surv_m_edu_stp_adj,
#   scale = 0.7,
#   height = 20,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# )

#Mother's Marital Status
surv_mar_status_stp_adj <- 
  ((all_klpn_plts[["m_mar_status"]] + labs(
    title = "Mother's Marital Status",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2a.m_mar_status"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2a.m_mar_status"]] + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  ))
  ) +
    plot_annotation(title = "Mother's Marital Status",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\surv_mar_status_stp_adj.png",
#   plot = surv_mar_status_stp_adj,
#   scale = 0.7,
#   height = 20,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# )

# All mother demography
surv_m_demo_all_stp_adj <-
  surv_m_age_stp_adj / surv_m_edu_stp_adj / surv_mar_status_stp_adj

ggsave(
  filename = "outputs\\surv_m_demo_all_stp_adj.png",
  plot = surv_m_demo_all_stp_adj,
  scale = 0.5,
  height = 70,
  width = 45,
  units = 'cm',
  dpi = 900
)
```

### Household Composition
```{r}
#Number of dependent children in HH
surv_hh_child_depndt_stp_adj <- 
  ((all_klpn_plts[["m_hh_child_depndt"]] + labs(
    title = "Number of Dependent Children in Household",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2a.m_hh_child_depndt"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2a.m_hh_child_depndt"]] + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  ))
  ) +
    plot_annotation(title = "Number of Dependent Children in Household",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\surv_hh_child_depndt_stp_adj.png",
#   plot = surv_hh_child_depndt_stp_adj,
#   scale = 0.7,
#   height = 20,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# )

#Number of Adults in HH
surv_hh_adult_stp_adj <- 
  ((all_klpn_plts[["m_hh_adult"]] + labs(
    title = "Number of Adults in Household",
    subtitle = "Unadjusted"
  )) + (cxph_adj_demo_plts[["demo2a.m_hh_adult"]] + labs(
    subtitle = "Demography adjusted",
    caption = NULL
  )) + (cxph_adj_full_plts[["full_2a.m_hh_adult"]] + labs(
    subtitle = "Fully adjusted",
    caption = NULL
  ))
  ) +
    plot_annotation(title = "Number of Adults in Household",
                    theme = theme(plot.title = element_text(size=14))
                    ) +
    plot_layout(guides = 'collect',
                axis_titles = "collect") &
    theme(legend.position='bottom')

# ggsave(
#   filename = "outputs\\surv_hh_adult_stp_adj.png",
#   plot = surv_hh_adult_stp_adj,
#   scale = 0.7,
#   height = 20,
#   width = 45,
#   units = 'cm',
#   dpi = 900
# )

# All HH Composition
surv_hh_comp_all_stp_adj <-
  surv_hh_child_depndt_stp_adj / surv_hh_adult_stp_adj

ggsave(
  filename = "outputs\\surv_hh_comp_all_stp_adj.png",
  plot = surv_hh_comp_all_stp_adj,
  scale = 0.5,
  height = 45,
  width = 45,
  units = 'cm',
  dpi = 900
)

save.image()
```

# Epilogue
